---
title: "R Notebook"
output: html_notebook
---

---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

Tools for working with COVID-19 Forecast Hub data: covidHubUtils R package
http://reichlab.io/covidHubUtils/articles/covidHubUtils-overview.html
```{r}
# devtools::install_github("reichlab/zoltr")
# devtools::install_github("reichlab/covidHubUtils")
```


```{r}
library(covidHubUtils)
library(doParallel)
library(tidyverse)

theme_set(theme_bw() + theme( legend.position="bottom" ) +
             theme( text = element_text(size=16, face="bold")) +
            theme(axis.text.x=element_text(angle=60, hjust=1)))

R1states <- c('CT', 'MA', 'ME', 'NH', 'RI', 'VT') # FEMA Region 1
R1stfips <- c('09', '25', '23', '33', '44', '50') # FIPS of states 
```


CDC ensemble forecasts for hospital admissions - https://github.com/reichlab/covid19-forecast-hub

```{r}
# Load forecasts that were submitted in a time window from zoltar
# inc_case_targets <- paste(1:4, "wk ahead inc case")
# forecasts_case <- load_forecasts(
#   models = "COVIDhub-ensemble",
#   dates = "2021-03-08",
#   date_window_size = 6,
#   locations = "US",
#   types = c("point", "quantile"),
#   targets = inc_case_targets,
#   source = "zoltar",
#   verbose = FALSE,
#   as_of = NULL,
#   hub = c("US")
# )

inc_hosp_targets <- paste(0:130, "day ahead inc hosp")
forecasts_hosp <- load_forecasts(
  models = "COVIDhub-ensemble",
  dates = "2022-01-17",
  date_window_size = 6,
  locations = R1stfips, #"US",
  types = c("point", "quantile"),
  targets = inc_hosp_targets,
  source = "zoltar",
  verbose = FALSE,
  as_of = NULL,
  hub = c("US")
)

# inc_death_targets <- paste(1:4, "wk ahead inc death")
# forecasts_death <- load_forecasts(
#   models = "COVIDhub-ensemble",
#   dates = "2021-03-08",
#   date_window_size = 6,
#   locations = "US",
#   types = c("point", "quantile"),
#   targets = inc_death_targets,
#   source = "zoltar",
#   verbose = FALSE,
#   as_of = NULL,
#   hub = c("US")
# )
```
HealthData Thruth
```{r}
truth_hosp <- load_truth(
  truth_source = "HealthData",
  target_variable = "inc hosp",
  locations = R1stfips # "US"
)
```
```{r}
df_plot <- bind_rows(
            truth_hosp
            , forecasts_hosp %>%
                filter(type == 'quantile') %>%
                pivot_wider(names_from = quantile, names_prefix = 'quant', values_from = value)
      ) %>%
  filter(target_end_date > "2021-11-01" #& !is.na(quant0.5)
         )
```

* POWER POINT Graph *

```{r, fig.height=4}
plotFcst <- function(tit, df_plot, ylims){
  brks <- sort(c(seq.Date(as.Date('2020-11-01'), as.Date('2022-02-01'), "1 month"),
                seq.Date(as.Date('2020-11-15'), as.Date('2022-02-01'), "1 month")))
ggplot(df_plot, aes(x = target_end_date)) +
    facet_wrap(~ abbreviation, ncol = 2, scales = 'free_y') + # - For different axes
    geom_point(aes(y = value, color = "Reported"), size = 1) +
    scale_color_manual(name = NULL, values = c("blue")) +
    geom_smooth(aes(y = value), color = "blue", size = 1, span = 0.1, se = FALSE) +
    geom_ribbon(aes(ymin = quant0.025, ymax = quant0.975, fill = "95% Prediction Interval"), size = 1) +
    geom_ribbon(aes(ymin = quant0.25, ymax = quant0.75, fill = "50% Prediction Interval"), size = 1) +
    scale_fill_manual(name = NULL, values = c("salmon3", "salmon1", "orangered4")) +
    geom_point(aes(y = quant0.5), color = "orangered4", size = 1) +
    geom_line(aes(y = quant0.5, fill = "Ensemble"), size = 1) +
    labs(x = NULL, y = 'New Hospital Admissions', title = tit) +
    expand_limits(y = ylims) +
    scale_x_date(breaks = brks, date_labels =  "%b %d %Y") +
    guides(color = guide_legend(order=1),
         fill = guide_legend(order=2))
}
```
```{r, fig.height=4}
plotFcst('Daily Hospital Admissions', df_plot, c(0, 100))
```
```{r, fig.height=4}
plotFcst('Daily Hospital Admissions per 100K people', df_plot %>%
           mutate(value = value / population * 1e5
                  , quant0.025 = quant0.025 / population * 1e5
                  , quant0.975 = quant0.975 / population * 1e5
                  , quant0.25 = quant0.25 / population * 1e5
                  , quant0.75 = quant0.75 / population * 1e5
                  , quant0.5 = quant0.5 / population * 1e5)
         , c(0, 13)) 
```

# MA ONLY
```{r, fig.height=2.5}
df_plot %>% 
  filter(abbreviation == 'MA') %>%
  ggplot(aes(x = target_end_date)) +
    geom_point(aes(y = value, color = "Reported"), size = 1) +
    scale_color_manual(name = NULL, values = c("blue")) +
    geom_smooth(aes(y = value), color = "blue", size = 1, span = 0.2, se = FALSE) +
    geom_ribbon(aes(ymin = quant0.025, ymax = quant0.975, fill = "95% Prediction Interval"), size = 1) +
    geom_ribbon(aes(ymin = quant0.25, ymax = quant0.75, fill = "50% Prediction Interval"), size = 1) +
    scale_fill_manual(name = NULL, values = c("salmon3", "salmon1", "orangered4")) +
    geom_point(aes(y = quant0.5), color = "orangered4", size = 1) +
    geom_line(aes(y = quant0.5, fill = "Ensemble"), size = 1) +
    labs(x = NULL, y = 'New Hospital Admissions', title = 'Massachusetts Daily Hospital Admission Forecast \nCDC Ensemble as of Jan 10th') +
    # expand_limits(y = ylims) +
    scale_x_date(breaks = sort(c(seq.Date(as.Date('2020-11-01'), as.Date('2022-01-01'), "1 month"),
                seq.Date(as.Date('2020-11-15'), as.Date('2022-01-15'), "1 month")))
                , date_labels =  "%b %d %Y") +
    guides(color = guide_legend(order=1),
         fill = guide_legend(order=2))
```

# FOR PPT - HealthData Thruth, daily hospitalizations nationwide
```{r}
truth_hosp_US <- load_truth(
  truth_source = "HealthData",
  target_variable = "inc hosp",
  locations = "US"
)
```
```{r}
truth_hosp_US %>%
  mutate(hospday100K = value / population * 1e5) %>%
  slice(tail(row_number(), 7)) %>%
  select(location, target_end_date, value, hospday100K) 
```
```{r}
truth_hosp_US %>%
  mutate(hospday100K = value / population * 1e5) %>%
  slice(tail(row_number(), 7)) %>%
  select(location, target_end_date, value, hospday100K) %>%
  summarise(hospday100K = mean(hospday100K))

```

# FOR PPT - JHU Thruth, daily incidence Nationwide
```{r}
truth_incidence_R1 <- load_truth(
  truth_source = "JHU",
  target_variable = "inc case",
  locations = R1stfips
)

truth_incidence_US <- load_truth(
  truth_source = "JHU",
  target_variable = "inc case",
  locations = "US"
)
```

```{r}
truth_incidence_US %>%
  mutate(incday100K = value / population * 1e5 / 7) %>%
  filter(target_end_date == max(target_end_date)) %>%
  select(location, target_end_date, value, incday100K)
```
# FOR PPT - JHU Thruth, daily deaths in R1
```{r}
truth_death <- load_truth(
  truth_source = "JHU",
  target_variable = "inc death",
  locations = R1stfips
)
```
```{r, fig.height=4}
truth_death %>%
  filter(target_end_date > '2020-09-01') %>%
  mutate(r = value/population * 1e5) %>%
ggplot(aes(x = target_end_date, y = r)) +
    facet_wrap(~ abbreviation, ncol = 2) + # , scales = 'free_y' - For different axes
    geom_point(color = "grey45", size = 1) +
    geom_smooth(color = "blue", size = 1, span = 0.1, se = FALSE) +
    labs(x = NULL, y = 'COVID-19 Deaths', title = 'COVID-19 Weekly Deaths per 100K people') +
    # expand_limits(y = ylims) +
    scale_x_date(breaks = seq.Date(as.Date('2020-11-01'), as.Date('2022-01-01'), "2 month"), date_labels =  "%b %d %Y")
```
```{r}
truth_death %>%
  filter(target_end_date > '2022-01-05') %>%
  mutate(r = value/population * 1e5) %>%
  select(target_end_date, location_name, r)
```
## DEATHS PER INCIDENCE
```{r}
tmp <- truth_incidence_R1 %>%
  group_by(location) %>%
  mutate(incidence_lagged = lag(value, n = 1)) %>%
  select(location, target_end_date, value, incidence_lagged)

death_rates <- truth_death %>%
  rename(deaths = value) %>%
  left_join(tmp, by = c('location', 'target_end_date')) %>%
  mutate(death_rate = deaths / incidence_lagged)
```

```{r, fig.height=4}
death_rates %>%
  filter(target_end_date > '2020-09-01') %>%
ggplot(aes(x = target_end_date, y = death_rate)) +
    facet_wrap(~ abbreviation, ncol = 2) + # , scales = 'free_y' - For different axes
    geom_point(color = "grey45", size = 1) +
    geom_smooth(color = "blue", size = 1, span = 0.1, se = FALSE) +
    labs(x = NULL, y = NULL, title = 'COVID-19 Weekly Deaths per New Cases (with a 2-week case lag)') +
    # expand_limits(y = ylims) +
    scale_x_date(breaks = seq.Date(as.Date('2020-11-01'), as.Date('2022-01-01'), "2 month"), date_labels =  "%b %d %Y") +
    scale_y_continuous(labels = scales::percent)
```
```{r, fig.height=4}
death_rates %>%
  filter(target_end_date > '2021-05-01') %>%
ggplot(aes(x = target_end_date, y = death_rate)) +
    facet_wrap(~ abbreviation, ncol = 2) + # , scales = 'free_y' - For different axes
    geom_point(color = "grey45", size = 1) +
    geom_smooth(color = "blue", size = 1, span = 0.1, se = FALSE) +
    labs(x = NULL, y = NULL, title = 'COVID-19 Weekly Deaths per New Cases (with a 2-week case lag)') +
    coord_cartesian(ylim = c(0, 0.04)) +
    scale_x_date(breaks = seq.Date(as.Date('2020-11-01'), as.Date('2022-01-01'), "2 month"), date_labels =  "%b %d %Y") +
    scale_y_continuous(labels = scales::percent)
```
