---
title: "R Notebook"
output: html_notebook
---

---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(zoo)
library(readxl)
library(lubridate)

theme_set(theme_bw() + theme( legend.position="bottom" ) +
             theme( text = element_text(size=16, face="bold")) +
            theme(axis.text.x=element_text(angle=60, hjust=1)))

# R1states <- c('CT', 'MA', 'ME', 'NH', 'RI', 'VT') # FEMA Region 1
# R1stfips <- c('09','25','23','33','44','50', '36') # FIPS of states in New England + NY
# R2states <- c('NY', 'NJ', 'PR')
# R3states <- c('DE', 'DC', 'MD', 'PA', 'VA', 'WV')
# R4states <- c('AL', 'FL', 'GA', 'KY', 'MS', 'NC', 'SC', 'TN') # Region 4
# R5states <- c('IL', 'IN', 'MI', 'MN', 'OH', 'WI')
# R6states <- c('TX', 'AR', 'LA', 'NM', 'OK')
# R7states <- c('IA', 'KS', 'MO', 'NE', '')
# R8states <- c('CO', 'MT', 'ND', 'SD', 'UT', 'WY')
# R9states <- c('AZ', 'CA', 'HI', 'NV') # Some Islands Missing Abriviations
# R10states <- c('AK', 'ID', 'OR', 'WA')

target_states <- c('DC', 'MS', 'NC', 'AK', 'NH', 'IL') # c('FL', 'MS', 'TX', 'CA') # Non-New England States to Compare
target_statefips <- c('11','28','37','02', '33', '17')
```


```{r}
df_population <- read_excel('data/NST-EST2021-POP.xlsx')
```

Clean up/Select Data
```{r}
df_population$state <- state.abb[match(df_population$`State Name`, state.name)]
df_population <- df_population %>%
  mutate(state = ifelse(`State Name` == "District of Columbia", "DC", state))

df_population <- df_population %>%
  mutate(region = case_when(
    # state %in% R1states ~ "Region 1",
    state %in% target_states ~ "Selected States"
    ))

df_population_r <- df_population %>%
  group_by(region) %>%
  summarise(
    Population = sum(Population, na.rm = TRUE)
    , .groups = 'keep'
  )  

```


```{r}
#hrrnums <- c(109,110,111,221,222,227,230,231,281,282,295,364,424)

# truth URL from the CDC ensemble IS THIS RAW OR MOVING AVG?
turl <- 'https://github.com/CSSEGISandData/COVID-19/raw/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv'
df_jhu <- read.csv( turl ) %>% tidyr::pivot_longer( cols = starts_with("X")
                                  , names_to="date", names_prefix="X"
                                  , values_to = "value",  values_drop_na = TRUE ) %>%
              mutate( date=as.Date(gsub("\\.","\\/",date), format="%m/%d/%y")
                      , location=sprintf("%05d", FIPS) ) %>%
              filter( substr(location, start=1, stop=2) %in% target_statefips )
```


```{r}
cases_by_state <- df_jhu %>%
  group_by(Province_State, date) %>%
  summarise(
    cases = sum(value, na.rm = TRUE)
    , .groups = 'keep'
  )

daily_cases_by_state <- cases_by_state %>%
  group_by(Province_State) %>%
  mutate(
    daily_cases = cases - lag(cases, default = first(cases), order_by = date)
    )

daily_cases_by_state_rates <- cases_by_state %>%
  group_by(Province_State) %>%
  mutate(
    daily_cases = cases - lag(cases, default = first(cases), order_by = date)
    )

daily_cases_by_state <- daily_cases_by_state %>%
  group_by(Province_State) %>%
  mutate(
    daily_cases = rollmean(daily_cases, k=7, fill = NA)
    )

daily_cases_by_state$state <- state.abb[match(daily_cases_by_state$Province_State, state.name)]
daily_cases_by_state <- daily_cases_by_state %>%
  mutate(state = ifelse(Province_State == "District of Columbia", "DC", state))

daily_cases_by_state <- rename(daily_cases_by_state, collection_date = date)
daily_cases_by_state <- daily_cases_by_state[ , -1]
```


load data from HHS Protect obtained by Jeff Cegan
```{r}
df_raw <- read.csv('data/[Unified] Hospital Analytic.csv') 
```

```{r}
facility <- df_raw %>%
  mutate(collection_date = as.Date(collection_date))
```



```{r}
summary(facility %>%
  select(hospital_pk, collection_date, total_beds, inpatient_beds, total_icu_beds)
  )
```



Aggregate by state
```{r}
state_d <- facility %>%
  group_by(state, collection_date) %>%
  summarise(
            total_all_beds = sum(total_beds, na.rm = TRUE)
            , inpatient_all_beds = sum(inpatient_beds, na.rm = TRUE)
            , icu_all_beds = sum(total_icu_beds, na.rm = TRUE)
            , total_child_beds = total_all_beds - sum(all_adult_hospital_beds, na.rm = TRUE)
            , inpatient_child_beds = inpatient_all_beds - sum(all_adult_hospital_inpatient_beds, na.rm = TRUE)
            , icu_child_beds = icu_all_beds - sum(total_staffed_adult_icu_beds, na.rm = TRUE)
            , inpatient_all_beds_used = sum(inpatient_beds_used, na.rm = TRUE)
            , inpatient_child_beds_used = inpatient_all_beds_used - sum(all_adult_hospital_inpatient_bed_occupied, na.rm = TRUE)
            , icu_all_beds_used = sum(icu_beds_used, na.rm = TRUE)
            , icu_child_beds_used = icu_all_beds_used - sum(staffed_adult_icu_bed_occupancy, na.rm = TRUE)
            , total_child_beds_used_covid_confirmed = sum(total_pediatric_patients_hospitalized_confirmed_covid, na.rm = TRUE)
            , total_child_beds_used_covid = sum(total_pediatric_patients_hospitalized_confirmed_and_suspected_covid, na.rm = TRUE)
            , total_all_beds_used_covid = total_child_beds_used_covid + sum(total_adult_patients_hospitalized_confirmed_and_suspected_covid, na.rm = TRUE)
            , admission_child_covid_confirmed = sum(previous_day_admission_pediatric_covid_confirmed, na.rm = TRUE)
            , admission_child_covid = admission_child_covid_confirmed +
                                            sum(previous_day_admission_pediatric_covid_suspected, na.rm = TRUE)
            , admission_all_covid = admission_child_covid 
                                        + sum(previous_day_admission_adult_covid_suspected 
                                              + previous_day_admission_adult_covid_confirmed
                                              , na.rm = TRUE)
            , admission_confirmed_covid = admission_child_covid_confirmed 
                                        + sum(previous_day_admission_adult_covid_confirmed, na.rm = TRUE)
            , total_utilization = inpatient_all_beds_used/inpatient_all_beds
            , covid_deaths = sum(deaths_covid, na.rm = TRUE)
            , staffing_shortage_yes = sum(critical_staffing_shortage_today == 'true', na.rm = TRUE)
            , total_hosp_reporting = n()
            , .groups = 'keep')

# moving average of the variables above
state_ma <- state_d[order(state_d$state, state_d$collection_date),] %>%
  group_by(state) %>%
  mutate(
            total_all_beds = rollmean(total_all_beds, k=7, fill=NA)
            , inpatient_all_beds = rollmean(inpatient_all_beds, k=7, fill=NA)
            , icu_all_beds = rollmean(icu_all_beds, k=7, fill=NA)
            , total_child_beds = rollmean(total_child_beds, k=7, fill=NA)
            , inpatient_child_beds = rollmean(inpatient_child_beds, k=7, fill=NA)
            , icu_child_beds = rollmean(icu_child_beds, k=7, fill=NA)
            , total_all_beds_used_covid = rollmean(total_all_beds_used_covid, k=7, fill=NA)
            , total_child_beds_used_covid_confirmed = rollmean(total_child_beds_used_covid_confirmed, k=7, fill=NA)
            , total_child_beds_used_covid = rollmean(total_child_beds_used_covid, k=7, fill=NA)
            , admission_all_covid = rollmean(admission_all_covid, k=7, fill=NA)
            , admission_child_covid_confirmed = rollmean(admission_child_covid_confirmed, k=7, fill=NA)
            , admission_child_covid = rollmean(admission_child_covid, k=7, fill=NA)
            , total_utilization = rollmean(total_utilization, k=7, fill=NA)
            , covid_deaths = rollmean(covid_deaths, k=7, fill=NA)
            ) 

state <- state_ma %>%
  pivot_longer(cols = -c(state, collection_date)
               , names_pattern = "(...*)_(...*)_beds", names_to = c("bedtype", "scope")) %>%
  mutate(bedtype=ifelse(bedtype=="", "value", bedtype)) 
# %>%  pivot_wider(names_from = 'scope', values_from = 'value')
```


Aggregate by Region
```{r}
region_d <- facility %>%
  group_by(fema_region, collection_date) %>%
  summarise(
            total_all_beds = sum(total_beds, na.rm = TRUE)
            , inpatient_all_beds = sum(inpatient_beds, na.rm = TRUE)
            , icu_all_beds = sum(total_icu_beds, na.rm = TRUE)
            , total_child_beds = total_all_beds - sum(all_adult_hospital_beds, na.rm = TRUE)
            , inpatient_child_beds = inpatient_all_beds - sum(all_adult_hospital_inpatient_beds, na.rm = TRUE)
            , icu_child_beds = icu_all_beds - sum(total_staffed_adult_icu_beds, na.rm = TRUE)
            , inpatient_all_beds_used = sum(inpatient_beds_used, na.rm = TRUE)
            , inpatient_child_beds_used = inpatient_all_beds_used - sum(all_adult_hospital_inpatient_bed_occupied, na.rm = TRUE)
            , icu_all_beds_used = sum(icu_beds_used, na.rm = TRUE)
            , icu_child_beds_used = icu_all_beds_used - sum(staffed_adult_icu_bed_occupancy, na.rm = TRUE)
            , total_child_beds_used_covid_confirmed = sum(total_pediatric_patients_hospitalized_confirmed_covid, na.rm = TRUE)
            , total_child_beds_used_covid = sum(total_pediatric_patients_hospitalized_confirmed_and_suspected_covid, na.rm = TRUE)
            , total_all_beds_used_covid = total_child_beds_used_covid + sum(total_adult_patients_hospitalized_confirmed_and_suspected_covid, na.rm = TRUE)
            , admission_child_covid_confirmed = sum(previous_day_admission_pediatric_covid_confirmed, na.rm = TRUE)
            , admission_child_covid = admission_child_covid_confirmed +
                                            sum(previous_day_admission_pediatric_covid_suspected, na.rm = TRUE)
            , admission_all_covid = admission_child_covid 
                                        + sum(previous_day_admission_adult_covid_suspected 
                                              + previous_day_admission_adult_covid_confirmed
                                              , na.rm = TRUE)
            , total_utilization = (inpatient_all_beds_used)/inpatient_all_beds
            , covid_deaths = sum(deaths_covid, na.rm = TRUE)
            , staffing_shortage_yes = sum(critical_staffing_shortage_today == 'yes', na.rm = TRUE)
            , total_hosp_reporting = n()
            , .groups = 'keep')

# moving average of the variables above
region_ma <- region_d[order(region_d$fema_region, region_d$collection_date),] %>%
  group_by(fema_region) %>%
  mutate(
            total_all_beds = rollmean(total_all_beds, k=7, fill=NA)
            , inpatient_all_beds = rollmean(inpatient_all_beds, k=7, fill=NA)
            , icu_all_beds = rollmean(icu_all_beds, k=7, fill=NA)
            , total_child_beds = rollmean(total_child_beds, k=7, fill=NA)
            , inpatient_child_beds = rollmean(inpatient_child_beds, k=7, fill=NA)
            , icu_child_beds = rollmean(icu_child_beds, k=7, fill=NA)
            , total_all_beds_used_covid = rollmean(total_all_beds_used_covid, k=7, fill=NA)
            , total_child_beds_used_covid_confirmed = rollmean(total_child_beds_used_covid_confirmed, k=7, fill=NA)
            , total_child_beds_used_covid = rollmean(total_child_beds_used_covid, k=7, fill=NA)
            , admission_all_covid = rollmean(admission_all_covid, k=7, fill=NA)
            , admission_child_covid_confirmed = rollmean(admission_child_covid_confirmed, k=7, fill=NA)
            , admission_child_covid = rollmean(admission_child_covid, k=7, fill=NA)
            , total_utilization = rollmean(total_utilization, k=7, fill=NA)
            , covid_deaths = rollmean(covid_deaths, k=7, fill=NA)
            ) 

region <- region_ma %>%
  pivot_longer(cols = -c(fema_region, collection_date)
               , names_pattern = "(...*)_(...*)_beds", names_to = c("bedtype", "scope")) %>%
  mutate(bedtype=ifelse(bedtype=="", "value", bedtype)) 
# %>%  pivot_wider(names_from = 'scope', values_from = 'value')
```

Hospitals
```{r}
state_d_h <- facility %>%
  group_by(hospital_name, collection_date, state) %>%
  summarise(
            total_all_beds = sum(total_beds, na.rm = TRUE)
            , inpatient_all_beds = sum(inpatient_beds, na.rm = TRUE)
            , icu_all_beds = sum(total_icu_beds, na.rm = TRUE)
            , total_child_beds = total_all_beds - sum(all_adult_hospital_beds, na.rm = TRUE)
            , inpatient_child_beds = inpatient_all_beds - sum(all_adult_hospital_inpatient_beds, na.rm = TRUE)
            , icu_child_beds = icu_all_beds - sum(total_staffed_adult_icu_beds, na.rm = TRUE)
            , inpatient_all_beds_used = sum(inpatient_beds_used, na.rm = TRUE)
            , inpatient_child_beds_used = inpatient_all_beds_used - sum(all_adult_hospital_inpatient_bed_occupied, na.rm = TRUE)
            , icu_all_beds_used = sum(icu_beds_used, na.rm = TRUE)
            , icu_child_beds_used = icu_all_beds_used - sum(staffed_adult_icu_bed_occupancy, na.rm = TRUE)
            , total_child_beds_used_covid_confirmed = sum(total_pediatric_patients_hospitalized_confirmed_covid, na.rm = TRUE)
            , total_child_beds_used_covid = sum(total_pediatric_patients_hospitalized_confirmed_and_suspected_covid, na.rm = TRUE)
            , total_all_beds_used_covid = total_child_beds_used_covid + sum(total_adult_patients_hospitalized_confirmed_and_suspected_covid, na.rm = TRUE)
            , admission_child_covid_confirmed = sum(previous_day_admission_pediatric_covid_confirmed, na.rm = TRUE)
            , admission_child_covid = admission_child_covid_confirmed +
                                            sum(previous_day_admission_pediatric_covid_suspected, na.rm = TRUE)
            , admission_all_covid = admission_child_covid 
                                        + sum(previous_day_admission_adult_covid_suspected 
                                              + previous_day_admission_adult_covid_confirmed
                                              , na.rm = TRUE)
            , non_child_covid_utilization = (inpatient_all_beds_used - total_child_beds_used_covid)/total_all_beds
            , non_covid_utilization = (inpatient_all_beds_used - total_all_beds_used_covid)/inpatient_all_beds
            , total_utilization = (inpatient_all_beds_used)/inpatient_all_beds
            , covid_deaths = sum(deaths_covid, na.rm = TRUE)
            , staffing_shortage_yes = sum(critical_staffing_shortage_today == 'yes', na.rm = TRUE)
            , .groups = 'keep')

# moving average of the variables above
state_ma_h <- state_d_h[order(state_d_h$hospital_name, state_d_h$collection_date, state_d_h$state),] %>%
  group_by(hospital_name) %>%
  mutate(
            total_all_beds = rollmean(total_all_beds, k=7, fill=NA)
            , inpatient_all_beds = rollmean(inpatient_all_beds, k=7, fill=NA)
            , icu_all_beds = rollmean(icu_all_beds, k=7, fill=NA)
            , total_child_beds = rollmean(total_child_beds, k=7, fill=NA)
            , inpatient_child_beds = rollmean(inpatient_child_beds, k=7, fill=NA)
            , icu_child_beds = rollmean(icu_child_beds, k=7, fill=NA)
            , total_all_beds_used_covid = rollmean(total_all_beds_used_covid, k=7, fill=NA)
            , total_child_beds_used_covid_confirmed = rollmean(total_child_beds_used_covid_confirmed, k=7, fill=NA)
            , total_child_beds_used_covid = rollmean(total_child_beds_used_covid, k=7, fill=NA)
            , admission_all_covid = rollmean(admission_all_covid, k=7, fill=NA)
            , admission_child_covid_confirmed = rollmean(admission_child_covid_confirmed, k=7, fill=NA)
            , admission_child_covid = rollmean(admission_child_covid, k=7, fill=NA)
            , non_child_covid_utilization = rollmean(non_child_covid_utilization, k=7, fill=NA)
            , non_covid_utilization = rollmean(non_covid_utilization, k=7, fill=NA)
            , total_utilization = rollmean(total_utilization, k=7, fill=NA)
            , covid_deaths = rollmean(covid_deaths, k=7, fill=NA)
            ) 
```


State Level Weekly Average Data for the Rates
```{r}
daily_cases_by_state_rates <- daily_cases_by_state_rates %>%
  group_by(Province_State) %>%
  mutate(
    week_year = paste(week(date), year(date), sep=" ")   
    )

daily_cases_by_state_rates$state <- state.abb[match(daily_cases_by_state_rates$Province_State, state.name)]
daily_cases_by_state_rates <- daily_cases_by_state_rates %>%
  mutate(state = ifelse(Province_State == "District of Columbia", "DC", state))

daily_cases_by_state_rates <- rename(daily_cases_by_state_rates, collection_date = date)
daily_cases_by_state_rates <- daily_cases_by_state_rates[ , -1]

state_ma_rates <- state_d[order(state_d$state, state_d$collection_date),] %>%
  group_by(state) %>%
  mutate(
    week_year = paste(week(collection_date), year(collection_date), sep=" ")
            ) 

daily_cases_hosp_rates_full <- state_ma_rates %>%
  full_join(daily_cases_by_state_rates, by = c("state","collection_date"))

daily_cases_hosp_rates_full <- daily_cases_hosp_rates_full %>%
  group_by(state, week_year.y) %>%
  summarise(
    weekly_cases = mean(daily_cases, na.rm = TRUE)
    , weekly_all_covid_admissions = mean(admission_all_covid, na.rm = TRUE)
    , weekly_confirmed_covid_admissions = mean(admission_confirmed_covid, na.rm = TRUE)
    , weekly_covid_deaths = mean(covid_deaths, na.rm = TRUE)
    , date = first(collection_date)
    , lag_cases = lag(weekly_cases, default = first(weekly_cases), k=2)
    , lag_cases_death = lag(weekly_cases, default = first(weekly_cases), k=4)
    , .groups = 'keep'
            ) 

```


Look at this year's staffing shortages
```{r}
temp <- facility %>%
  filter(state %in% target_states & collection_date == '2021-12-29')

table(temp$critical_staffing_shortage_today)
table(temp$critical_staffing_shortage_anticipated_within_week)

crit_shortage_hosp <- temp %>%
  filter(critical_staffing_shortage_today == 'true')

crit_shortage_hosp_week <- temp %>%
  filter(critical_staffing_shortage_anticipated_within_week == 'true')

crit_shortage_hosp_day_or_week <- temp %>%
  filter(critical_staffing_shortage_anticipated_within_week == 'true' | critical_staffing_shortage_today == 'true')

```

# SLide Graph - Number of hospitals per state reporting lately
```{r, fig.height=4}
state_d %>%
  filter( state %in% target_states & collection_date > '2022-01-13') %>% 
  mutate(r = staffing_shortage_yes/total_hosp_reporting) %>%
  select(state, collection_date, r, total_hosp_reporting)
```
Staffing Shortage Graph
```{r, fig.height=4}
# IMPORTANT
date_most_hospitals_reported <- '2022-01-18'

state_d %>%
  # filter( state %in% target_states & total_hosp_reporting > 5 & collection_date > '2020-09-01') %>% 
  filter( state %in% target_states & collection_date > '2020-09-01' & collection_date <= date_most_hospitals_reported) %>% 
  mutate(r = staffing_shortage_yes/total_hosp_reporting) %>%
  ggplot(aes(x = collection_date, y = r)) + #, color = state
    facet_wrap(~ state, ncol = 2) + #scales = 'free_y' - For different axes
    geom_point(color = "grey55", size = 1) +
    geom_smooth(color = "blue", size = 1, span = 0.2) +
    labs(x = NULL, y = 'Percent of Hospitals Reporting Shortages', title = 'Hospital Staffing Shortages') +
    # scale_x_date(date_breaks = "2 month", date_labels =  "%b-%d-%Y") +
    scale_x_date(breaks = seq.Date(as.Date('2020-11-01'), as.Date('2022-01-01'), "2 month"), date_labels =  "%b %d %Y") +
    scale_y_continuous(labels = scales::percent)
```
```{r}
state_d %>%
  filter( state == "DC" & collection_date > '2021-12-01') %>% 
  mutate(r = staffing_shortage_yes/total_hosp_reporting) %>%
  select(state, collection_date, r, staffing_shortage_yes, total_hosp_reporting)
```


Join HHS and Census Data
```{r}
state_ma <- state_ma %>%
  full_join(df_population, by = "state")
```

# * POWER POINT Graph *
Hospital Admissions per 100,000 Ppl
```{r, fig.height=4}
state_ma %>%
  filter( state %in% target_states & collection_date > '2020-09-01' & collection_date <= date_most_hospitals_reported) %>%
  mutate(r = (admission_all_covid/Population)*100000) %>%
  ggplot(aes(x = collection_date, y = r)) + #, color = state
    facet_wrap(~ state, ncol = 2) + #scales = 'free_y' - For different axes
    geom_line(color = "blue", size = 1) +
    labs(x = NULL, y = 'COVID-19 Hospitalizations', title = 'COVID-19 Daily Hospital Admissions per 100K People,\n 7-day moving average') +
    # scale_x_date(date_breaks = "2 month", date_labels =  "%b-%d-%Y")
    scale_x_date(breaks = seq.Date(as.Date('2020-11-01'), as.Date('2022-01-01'), "2 month"), date_labels = "%b %d %Y")
```
## Slide - R1 Average of Hosp admissions
```{r}
state_ma %>%
  filter( state %in% target_states & collection_date > '2022-01-13') %>%
  ungroup() %>% group_by(collection_date) %>%
  summarise(r = sum(admission_all_covid) / sum(Population) * 1e5)
```


Join HHS and Case Data
```{r}
daily_cases_hosp <- state_ma %>%
  full_join(daily_cases_by_state, by = c("state","collection_date"))
```

# SLide = Hospital Utilization per State
```{r, fig.height=4}
daily_cases_hosp %>%
  filter(state %in% target_states & collection_date > '2020-09-15') %>%
  ggplot(aes(x = collection_date, y = total_utilization)) + 
    facet_wrap(~ state, ncol = 2) +
    geom_hline(yintercept = 0.8, color = 'orange', linetype='dashed') +
    geom_point(color = "gray55", size = 1) +
    geom_smooth(color = "blue", size = 1, span = 0.1, se = FALSE) +
    labs(x = NULL, y = NULL, title = 'Total Hospital Utilization, 7-day Moving Average') +
    scale_x_date(breaks = seq.Date(as.Date('2020-11-01'), as.Date('2022-01-01'), "2 month"), date_labels =  "%b %d %Y") +
    scale_y_continuous(labels=scales::percent, limits = c(0, 1))
```
  
# Slide
Case Incidence (all, not just hospitalizations) Per 100,000 People
```{r, fig.height=4}
daily_cases_hosp %>%
  filter(state %in% target_states & collection_date > '2020-09-15') %>% 
  mutate(r = (daily_cases/Population)*100000) %>%
  ggplot(aes(x = collection_date, y = r)) + #, color = state
    facet_wrap(~ state, ncol = 2) + #scales = 'free_y' - For different axes
    # geom_line(color = "blue", size = 1) +
    geom_point(color = "gray45", size = 1) +
    geom_smooth(color = "blue", size = 1, span = 0.02, se = FALSE) +
    labs(x = NULL, y = 'COVID-19 Cases', title = 'COVID-19 Daily Cases per 100K People, Weekly Average') +
    scale_x_date(breaks = seq.Date(as.Date('2020-11-01'), as.Date('2022-01-01'), "2 month"), date_labels =  "%b %d %Y")
```

# SLIDE: Resource Prioritization per State - REVIEW SHORTAGE DATE BELOW
```{r}
state_summary <- daily_cases_hosp %>%
  filter(state %in% target_states) %>%
  filter(!is.na(total_utilization)) %>%
  group_by(state) %>%
  mutate(maxdate = max(collection_date, na.rm = TRUE)) 
#if (length(unique(state_summary$maxdate)) > 1){
  state_summary$maxdate = date_most_hospitals_reported # min(unique(state_summary$maxdate))
  state_summary$maxdate = '2022-01-17' # min(unique(state_summary$maxdate))
#}
cols <- c('state', paste('Hospital Utilization as of', unique(state_summary$maxdate)))
state_summary <- state_summary %>%
  filter(collection_date == maxdate) %>%
  select(state, total_utilization)
colnames(state_summary) <- cols
  
tmp <- daily_cases_hosp %>%
  filter(state %in% target_states) %>%
  mutate(inc100k = (daily_cases/Population)*100000) %>%
  filter(!is.na(inc100k)) %>%
  group_by(state) %>%
  mutate(maxdate = max(collection_date, na.rm = TRUE)) 
if (length(unique(tmp$maxdate)) > 1){
  tmp$maxdate = min(unique(tmp$maxdate))
}
cols <- c('state', paste('New Cases per 100K as of', unique(tmp$maxdate)))
tmp <- tmp %>%
  filter(collection_date == maxdate) %>%
  select(state, inc100k)
colnames(tmp) <- cols

state_summary <- state_summary %>%
  left_join(tmp, by = c('state'))
  
tmp <- state_ma %>%
  filter(state %in% target_states) %>%
  mutate(r = (admission_all_covid/Population)*100000) %>%
  filter(!is.na(r)) %>%
  group_by(state) %>%
  mutate(maxdate = max(collection_date, na.rm = TRUE)) 
if (length(unique(tmp$maxdate)) > 1){
  tmp$maxdate = min(unique(tmp$maxdate))
}
cols <- c('state', paste('Hosp. Admissions per 100K as of', unique(tmp$maxdate)))
tmp <- tmp %>%
  filter(collection_date == maxdate) %>%
  select(state, r)
colnames(tmp) <- cols

state_summary <- state_summary %>%
  left_join(tmp, by = c('state'))

tmp <- state_d %>%
  filter( state %in% target_states) %>%
  mutate(r = staffing_shortage_yes/total_hosp_reporting) %>%
  filter(total_hosp_reporting > 5 & !is.na(r)) %>%
  group_by(state) %>%
  filter((state != 'MA' & collection_date == date_most_hospitals_reported)
         | (state == 'MA'& collection_date == '2022-01-09')) %>%
  select(state, r, collection_date)
colnames(tmp) <- c('state', 'Staffing Shortage', 'Shortage as of')

state_summary <- state_summary %>%
  left_join(tmp, by = c('state')) #%>%
  #arrange(match(state, c("RI", "MA", "CT", "NH", "VT", "ME")))
```


```{r}
state_summary %>%
  kableExtra::kbl() %>% kableExtra::kable_styling()
```


COVID admissions are confirmed ONLY
Hospitalization Rate
```{r, fig.height=4}
daily_cases_hosp_rates_full %>%
  filter(state %in% target_states & date > '2020-09-01') %>%
  mutate(r = weekly_confirmed_covid_admissions/lag_cases) %>%
  filter(r < 2) %>%
  ggplot(aes(x = date, y = r)) + #, color = state
    facet_wrap(~ state, ncol = 2) + #scales = 'free_y' - For different axes
    # geom_line(color = "blue", size = 1) +
    geom_point(color = "gray45", size = 1) +
    geom_smooth(color = "blue", size = 1, span = 0.1, se = FALSE) +
    labs(x = NULL, y = NULL, title = 'Confirmed COVID-19 Hospitalization Rate,\n 14-day moving average with a 2 week case lag') +
    scale_x_date(breaks = seq.Date(as.Date('2020-11-01'), as.Date('2022-01-01'), "2 month"), date_labels =  "%b %d %Y") +
    scale_y_continuous(labels = scales::percent)
```
```{r}
state_ma_rates %>%
  filter(state %in% target_states ) %>%
  group_by(state) %>%
  summarise(n = n())
```

```{r, fig.height=4}
daily_cases_hosp_rates_full %>%
  filter(state %in% target_states & date > '2021-05-01') %>%
  mutate(r = weekly_confirmed_covid_admissions/lag_cases) %>%
  filter(r < 2) %>%
  ggplot(aes(x = date, y = r)) + #, color = state
    facet_wrap(~ state, ncol = 2) + #scales = 'free_y' - For different axes
    # geom_line(color = "blue", size = 1) +
    geom_point(color = "gray45", size = 1) +
    geom_smooth(color = "blue", size = 1, span = 0.1, se = FALSE) +
    labs(x = NULL, y = NULL, title = 'Confirmed COVID-19 Hospitalization Rate,\n 14-day moving average with a 2 week case lag') +
    scale_x_date(breaks = seq.Date(as.Date('2020-11-01'), as.Date('2022-01-01'), "2 month"), date_labels =  "%b %d %Y") +
    scale_y_continuous(labels = scales::percent) +
    coord_cartesian(ylim = c(0, 0.15))
```

```{r}
daily_cases_hosp_rates_full %>%
  filter(state %in% target_states & date > '2022-01-01' ) %>% #& date < '2021-10-01'
  mutate(r = (weekly_confirmed_covid_admissions/lag_cases)*100) %>%
  filter(r < 200) %>%
  group_by(state) %>%
  summarise(
    mean_rate = mean(r)
  )
```



Join HHS and Census Data
```{r}
region_ma <- region_ma %>%
  full_join(df_population_r, by = c("fema_region" = "region"))
```

* POWER POINT Stat*
Admissions Incidence per 100,000 People, Regionally
```{r, fig.height=4}
region_ma %>%
  filter(fema_region %in% c("Region 1") & collection_date > '2022-01-11') %>% #& collection_date < '2020-12-25'
  mutate(r = (admission_all_covid/Population)* 1e5) %>%
  select(fema_region, collection_date, admission_all_covid, inpatient_all_beds, r)

```


Aggregate by Date
```{r}
us_d <- facility %>%
  group_by(collection_date) %>%
  summarise(
            total_child_beds_used_covid = sum(total_pediatric_patients_hospitalized_confirmed_and_suspected_covid, na.rm = TRUE)
            , total_all_beds_used_covid = total_child_beds_used_covid + sum(total_adult_patients_hospitalized_confirmed_and_suspected_covid, na.rm = TRUE)
            , admission_child_covid_confirmed = sum(previous_day_admission_pediatric_covid_confirmed, na.rm = TRUE)
            , admission_child_covid = admission_child_covid_confirmed +
                                            sum(previous_day_admission_pediatric_covid_suspected, na.rm = TRUE)
            , admission_all_covid = admission_child_covid 
                                        + sum(previous_day_admission_adult_covid_suspected 
                                              + previous_day_admission_adult_covid_confirmed
                                              , na.rm = TRUE)
            , covid_deaths = sum(deaths_covid, na.rm = TRUE)
            , .groups = 'keep')

# moving average of the variables above
us_ma <- us_d[order(us_d$collection_date),] %>%
  ungroup() %>%
  mutate(
            total_all_beds_used_covid = rollmean(total_all_beds_used_covid, k=7, fill=NA)
            , total_child_beds_used_covid = rollmean(total_child_beds_used_covid, k=7, fill=NA)
            , admission_all_covid = rollmean(admission_all_covid, k=7, fill=NA)
            , admission_child_covid_confirmed = rollmean(admission_child_covid_confirmed, k=7, fill=NA)
            , admission_child_covid = rollmean(admission_child_covid, k=7, fill=NA)
            , covid_deaths = rollmean(covid_deaths, k=7, fill=NA)
            ) 

```

Admissions per 100,000
```{r}
us_ma %>%
  filter(collection_date > '2022-01-11') %>%
  mutate(r = (admission_all_covid / 329500000)* 1e5) %>%
  select(collection_date, admission_all_covid, r)
```



Hospital Utilization By State
```{r, fig.height = 4, fig.width = 4}
state_ma %>%
  filter(state %in% target_states & collection_date > '2022-01-11') %>%
  select(
    state
    , collection_date
    , total_utilization
  )
```


Hospital Utilization by Hospital
```{r}
RI_hosp_utilization <- state_ma_h %>%
  filter(state %in% c("RI") & collection_date == '2022-01-03') %>%
  select(
    state
    , hospital_name
    , collection_date
    , total_utilization
    , non_covid_utilization
  )


```

Save as Excel
```{r}
fname1 <- paste0('2022-01-03_RI_hosp_util.xlsx')

openxlsx::write.xlsx( 
              RI_hosp_utilization
              , fname1 )
```



Review Specific Hospital's Utilization
```{r, fig.height=4}
# Hospitals: 
ri_hosp <- c("KENT COUNTY MEMORIAL HOSPITAL", "LANDMARK MEDICAL CENTER", "RHODE ISLAND HOSPITAL", "THE MIRIAM HOSPITAL")


state_ma_h %>% 
  filter( hospital_name %in% ri_hosp & collection_date > '2020-09-01') %>%
  mutate(r1 = total_utilization*100) %>%
  mutate(r2 = non_covid_utilization*100) %>%
  ggplot() +  
  facet_wrap(~ hospital_name, ncol = 2) +
  geom_ribbon(aes(ymin = r2, ymax = r1, x=collection_date), fill = "blue") + 
  geom_line(aes(x = collection_date, y = r1, color = "Total Utilization")) +
  geom_line(aes(x = collection_date, y = r2, color = "Non-COVID-19 Utilization")) +
  labs(title = "Hospital Utilization, 7-day moving average", color = "") +
  ylim(0, 100) +
  scale_x_date(date_breaks = "2 month", date_labels =  "%b-%d-%Y") +
  ylab('Percent (%)')
```

```{r, fig.height=4}
# Hospitals: 
ri_hosp2 <- c("OUR LADY OF FATIMA HOSPITAL", "NEWPORT HOSPITAL", "WESTERLY HOSPITAL", "WOMEN & INFANTS HOSPITAL OF RHODE ISLAND", "ROGER WILLIAMS MEDICAL CENTER", "SOUTH COUNTY HOSPITAL INC")


state_ma_h %>% 
  filter( hospital_name %in% ri_hosp2 & collection_date > '2020-09-01') %>%
  mutate(r1 = total_utilization*100) %>%
  mutate(r2 = non_covid_utilization*100) %>%
  filter(r2 > 0) %>%
  ggplot() +  
  facet_wrap(~ hospital_name, ncol = 2) +
  geom_ribbon(aes(ymin = r2, ymax = r1, x=collection_date), fill = "blue") + 
  geom_line(aes(x = collection_date, y = r1, color = "Total Utilization")) +
  geom_line(aes(x = collection_date, y = r2, color = "Non-COVID-19 Utilization")) +
  labs(title = "Hospital Utilization, 7-day moving average", color = "") +
  ylim(0, 100) +
  scale_x_date(date_breaks = "2 month", date_labels =  "%b-%d-%Y") +
  ylab('Percent (%)')
```





See the latest report date and review col's when needed
```{r}
max(facility$collection_date)
#colnames(facility)
#temp$staffing_shortage_details
```



Code that is being retired
-----------------------------------------------
Staffing Shortage is a Graph Now not a table:
Look at last year's staffing shortages
```{r}
temp_prev <- facility %>%
  filter(state %in% target_states & collection_date == '2020-12-09')

table(temp_prev$critical_staffing_shortage_today)
table(temp_prev$critical_staffing_shortage_anticipated_within_week)

crit_shortage_hosp_prev <- temp_prev %>%
  filter(critical_staffing_shortage_today == 'true')

crit_shortage_hosp_week_prev <- temp_prev %>%
  filter(critical_staffing_shortage_anticipated_within_week == 'true')

crit_shortage_hosp_day_or_week_prev <- temp_prev %>%
  filter(critical_staffing_shortage_anticipated_within_week == 'true' | critical_staffing_shortage_today == 'true')

inner_join(crit_shortage_hosp_prev, crit_shortage_hosp_week_prev, by = "ccn")
```

Look at this year's staffing shortages
```{r}
temp <- facility %>%
  filter(state %in% target_states & collection_date == '2021-12-22')

table(temp$critical_staffing_shortage_today)
table(temp$critical_staffing_shortage_anticipated_within_week)

crit_shortage_hosp <- temp %>%
  filter(critical_staffing_shortage_today == 'true')

crit_shortage_hosp_week <- temp %>%
  filter(critical_staffing_shortage_anticipated_within_week == 'true')

crit_shortage_hosp_day_or_week <- temp %>%
  filter(critical_staffing_shortage_anticipated_within_week == 'true' | critical_staffing_shortage_today == 'true')

inner_join(crit_shortage_hosp, crit_shortage_hosp_week, by = "ccn")
```

See the common hospitals who have and are expecting shortages
```{r}
common_hosp <- inner_join(crit_shortage_hosp_day_or_week, crit_shortage_hosp_day_or_week_prev, by = "ccn")
#inner_join(crit_shortage_hosp_day_or_week, crit_shortage_hosp_day_or_week_prev, by = "ccn")
#inner_join(crit_shortage_hosp_day_or_week, crit_shortage_hosp_day_or_week_prev, by = "hospital_name")
```

Save Excel Sheet of staffing shortages
```{r}
fname1 <- paste0('2021-12-15_staff_shortage.xlsx')

openxlsx::write.xlsx( 
              crit_shortage_hosp_day_or_week %>% select(hospital_name, state, county, zip, critical_staffing_shortage_today, critical_staffing_shortage_anticipated_within_week)
              , fname1 )

```

State Loop! For staffing shortages (track by state)
```{r}
for (state_abbrv in target_states){
  tempState <- facility %>%
  filter(state %in% state_abbrv & collection_date == '2021-12-15')

  print(state_abbrv)
  print(table(tempState$critical_staffing_shortage_today))
  print(table(tempState$critical_staffing_shortage_anticipated_within_week))

  crit_shortage_hospState <- tempState %>%
    filter(critical_staffing_shortage_today == 'true')

  crit_shortage_hosp_weekState <- tempState %>%
    filter(critical_staffing_shortage_anticipated_within_week == 'true')

  crit_shortage_hosp_day_or_weekState <- tempState %>%
    filter(critical_staffing_shortage_anticipated_within_week == 'true' | critical_staffing_shortage_today == 'true')

  print(inner_join(crit_shortage_hospState, crit_shortage_hosp_weekState, by = "ccn"))
  
}

```




NOT USING THIS SOURCE ANYMORE:
Deaths
```{r, fig.height=4}
daily_cases_hosp %>%
  filter(state %in% target_states & collection_date > '2020-09-01') %>%
  mutate(r = (covid_deaths/Population)*100000) %>%
  ggplot(aes(x = collection_date, y = r)) + #, color = state
    facet_wrap(~ state, ncol = 2) + #scales = 'free_y' - For different axes
    geom_line(color = "blue", size = 1) +
    labs(x = NULL, y = 'COVID-19 Deaths', title = 'COVID-19 Daily Deaths per 100,000 People,\n 7-day moving average') +
    scale_x_date(date_breaks = "2 month", date_labels =  "%b-%d-%Y") 
```

NOT USING THIS SOURCE ANYMORE
Death Rate
```{r, fig.height=4}
daily_cases_hosp_rates_full %>%
  filter(state %in% target_states & date > '2020-09-01') %>%
  mutate(r = weekly_covid_deaths/lag_cases_death) %>%
  ggplot(aes(x = date, y = r)) + #, color = state
    facet_wrap(~ state, ncol = 2) + #scales = 'free_y' - For different axes
    geom_line(color = "blue", size = 1) +
    labs(x = NULL, y = 'COVID-19 Deaths/Cases', title = 'COVID-19 Death Rate,\n Weely Average with a 4 week case lag') +
    scale_x_date(date_breaks = "2 month", date_labels =  "%b-%d-%Y") 
```


```{r}
daily_cases_hosp_rates_full %>%
  filter(state %in% target_states & date > '2021-12-01') %>%
  mutate(r = (weekly_covid_deaths/lag_cases_death)) %>%
  group_by(state) %>%
  summarise(
    mean_rate = mean(r)
  )
```


Staffing shortages at the hospital level for RI
```{r}
temp <- facility %>%
  filter(state %in% c("RI") & collection_date == '2022-01-05')

table(temp$critical_staffing_shortage_today)
table(temp$critical_staffing_shortage_anticipated_within_week)

crit_shortage_hosp <- temp %>%
  filter(critical_staffing_shortage_today == 'true')

crit_shortage_hosp_week <- temp %>%
  filter(critical_staffing_shortage_anticipated_within_week == 'true')

crit_shortage_hosp_day_or_week <- temp %>%
  filter(critical_staffing_shortage_anticipated_within_week == 'true' | critical_staffing_shortage_today == 'true')

```

```{r}
fname1 <- paste0('2022-01-05_RI_staff_shortage.xlsx')

openxlsx::write.xlsx( 
              temp %>% select(hospital_name, state, county, zip, hospital_subtype, critical_staffing_shortage_today, critical_staffing_shortage_anticipated_within_week)
              , fname1 )
```


```{r}
temp %>% 
  select(hospital_name
         , state
         , county
         , zip
         , hospital_subtype
         , system
         , critical_staffing_shortage_today
         , critical_staffing_shortage_anticipated_within_week)
```

Checking Deaths in CT
```{r}
sanity_check <- facility %>%
  filter(state == 'CT') %>%
  summarise(
    #covid_deaths = sum(deaths_covid, na.rm = TRUE)
    deaths_covid
    , collection_date
    , hospital_name
  )
```

```{r}
daily_cases_hosp %>%
  mutate(inc100k = (daily_cases/Population)*100000) %>%
  filter(!is.na(inc100k) & state == 'RI' & collection_date > '2021-12-21') %>%
  select(state, collection_date, inc100k)
```

COVID admissions are confirmed and suspected
Hospitalization Rate - LOOKS INCONSISTENT - we are current using confirmed admissions only
```{r, fig.height=4}
daily_cases_hosp_rates_full %>%
  filter(state %in% target_states & date > '2020-09-01') %>%
  mutate(r = (weekly_all_covid_admissions/lag_cases)*100) %>%
  filter(r < 200) %>%
  ggplot(aes(x = date, y = r)) + #, color = state
    facet_wrap(~ state, ncol = 2) + #scales = 'free_y' - For different axes
    geom_line(color = "blue", size = 1) +
    labs(x = NULL, y = 'Percent', title = 'Confirmed and Suspected COVID-19 Hospitalization Rate,\n Weely Average with a 2 week case lag') +
    scale_x_date(date_breaks = "2 month", date_labels =  "%b-%d-%Y") 
```