---
title: "HHS Presentation Combined"
author: "Luke Hogewood"
date: "2/2/2022"
output: html_document
---
Consolidated code from Alex Ligo and Cat Sherman

Tools for working with COVID-19 Forecast Hub data: covidHubUtils R package
http://reichlab.io/covidHubUtils/articles/covidHubUtils-overview.html
```{r}
# BEFORE RUNNING THIS NOTEBOOK, UNCOMMENT THIS CELL TO RUN THE COMMANDS BELOW
# devtools::install_github("reichlab/zoltr")
# devtools::install_github("reichlab/covidHubUtils")
```

# Import Libraries
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(zoo)
library(readxl)
library(lubridate)
library(covidHubUtils)
library(doParallel)
library(tidyverse)

theme_set(theme_bw() + theme( legend.position="bottom" ) +
             theme( text = element_text(size=16, face="bold")) +
            theme(axis.text.x=element_text(angle=60, hjust=1)))

R1states <- c('CT', 'MA', 'ME', 'NH', 'RI', 'VT') # FEMA Region 1
R1stfips <- c('09','25','23','33','44','50') #, '36') # FIPS of states in New England + NY
#R2states <- c('NY', 'NJ', 'PR')
#R3states <- c('DE', 'DC', 'MD', 'PA', 'VA', 'WV')
#R4states <- c('AL', 'FL', 'GA', 'KY', 'MS', 'NC', 'SC', 'TN') # Region 4
#R5states <- c('IL', 'IN', 'MI', 'MN', 'OH', 'WI')
#R6states <- c('TX', 'AR', 'LA', 'NM', 'OK')
#R7states <- c('IA', 'KS', 'MO', 'NE', '')
#R8states <- c('CO', 'MT', 'ND', 'SD', 'UT', 'WY')
#R9states <- c('AZ', 'CA', 'HI', 'NV') # Some Islands Missing Abbreviations
#R10states <- c('AK', 'ID', 'OR', 'WA')

#Selectstates <- c('FL', 'MS', 'TX', 'CA') # Non-New England States to Compare

target_states <- R1states
target_statefips <- R1stfips
```


# Set Important Variables
```{r}
# Date most hospitals reported for HHS dataset
# Typically this is the Wednesday of the previous week
date_most_hospitals_reported <- '2022-02-09'

# Date all states report (issues w/ RI): 1 day before date most hospitals reported
state_reported <- '2022-02-08'

# CDC Forecast Date: Date of most recent CDC ensemble updates
# Typically this is every Monday night
date_latest_forecast <- '2022-02-14'

# JHU forecast date: usually previous Friday
date_case_count <- '2022-02-11'

# Test date: more recent date to see more granular view
test_date <- '2021-10-01'
```


# Download Datasets

## CDC Ensemble Forecast
```{r}
inc_hosp_targets <- paste(0:130, "day ahead inc hosp")
forecasts_hosp <- load_forecasts(
  models = "COVIDhub-ensemble",
  dates = date_latest_forecast,
  date_window_size = 6,
  locations = target_statefips, #"US",
  types = c("point", "quantile"),
  targets = inc_hosp_targets,
  source = "zoltar",
  verbose = FALSE,
  as_of = NULL,
  hub = c("US")
)
```


## Population Breakdown
```{r}
df_population <- read_excel('Import data/NST-EST2021-POP.xlsx')

df_population$state <- state.abb[match(df_population$`State Name`, state.name)]
df_population <- df_population %>%
  mutate(state = ifelse(`State Name` == "District of Columbia", "DC", state))

df_population <- df_population %>%
  mutate(region = case_when(
    # state %in% R1states ~ "Region 1",
    state %in% target_states ~ "Selected States"
    ))

df_population_r <- df_population %>%
  group_by(region) %>%
  summarise(
    Population = sum(Population, na.rm = TRUE)
    , .groups = 'keep'
  )  
```

## Johns Hopkins COVID Case Data
```{r}
# truth URL from the CDC ensemble IS THIS RAW OR MOVING AVG?
turl <- 'https://github.com/CSSEGISandData/COVID-19/raw/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv'
df_jhu <- read.csv( turl ) %>% tidyr::pivot_longer( cols = starts_with("X")
                                  , names_to="date", names_prefix="X"
                                  , values_to = "value",  values_drop_na = TRUE ) %>%
              mutate( date=as.Date(gsub("\\.","\\/",date), format="%m/%d/%y")
                      , location=sprintf("%05d", FIPS) ) %>%
              filter( substr(location, start=1, stop=2) %in% target_statefips )

# Wrangle JHU Case data
cases_by_state <- df_jhu %>%
  group_by(Province_State, date) %>%
  summarise(
    cases = sum(value, na.rm = TRUE)
    , .groups = 'keep'
  )

daily_cases_by_state <- cases_by_state %>%
  group_by(Province_State) %>%
  mutate(
    daily_cases = cases - lag(cases, default = first(cases), order_by = date)
    )

daily_cases_by_state_rates <- cases_by_state %>%
  group_by(Province_State) %>%
  mutate(
    daily_cases = cases - lag(cases, default = first(cases), order_by = date)
    )

daily_cases_by_state <- daily_cases_by_state %>%
  group_by(Province_State) %>%
  mutate(
    daily_cases = rollmean(daily_cases, k=7, fill = NA)
    )

daily_cases_by_state$state <- state.abb[match(daily_cases_by_state$Province_State, state.name)]
daily_cases_by_state <- daily_cases_by_state %>%
  mutate(state = ifelse(Province_State == "District of Columbia", "DC", state)) %>%
  filter(state %in% target_states)

daily_cases_by_state <- rename(daily_cases_by_state, collection_date = date)
daily_cases_by_state <- daily_cases_by_state[ , -1]
```

## HHS Protect data obtained by Jeff Cegan
```{r}
df_raw <- read.csv('Import data/[Unified] Hospital Analytic.csv') %>%
    filter(state %in% c('CT', 'MA', 'ME', 'NH', 'RI', 'VT', 'NY')) %>%
		mutate(collection_date = as.Date(collection_date)) %>%
		filter(collection_date > '2020-09-30')
colnames(df_raw)[1] <- gsub('^...','',colnames(df_raw)[1])

summary(df_raw %>%
  select(hospital_pk, collection_date, total_beds, inpatient_beds, total_icu_beds)
  )
```

# State-level Analysis

## Group and Summarize by State
```{r}

state_d <- df_raw %>%
  filter(state %in% target_states) %>%
  group_by(state, collection_date) %>%
  summarise(
            total_all_beds = sum(total_beds, na.rm = TRUE)
            , inpatient_all_beds = sum(inpatient_beds, na.rm = TRUE)
            , icu_all_beds = sum(total_icu_beds, na.rm = TRUE)
            , total_child_beds = total_all_beds - sum(all_adult_hospital_beds, na.rm = TRUE)
            , inpatient_child_beds = inpatient_all_beds - sum(all_adult_hospital_inpatient_beds, na.rm = TRUE)
            , icu_child_beds = icu_all_beds - sum(total_staffed_adult_icu_beds, na.rm = TRUE)
            , inpatient_all_beds_used = sum(inpatient_beds_used, na.rm = TRUE)
            , inpatient_child_beds_used = inpatient_all_beds_used - sum(all_adult_hospital_inpatient_bed_occupied, na.rm = TRUE)
            , icu_all_beds_used = sum(icu_beds_used, na.rm = TRUE)
            , icu_child_beds_used = icu_all_beds_used - sum(staffed_adult_icu_bed_occupancy, na.rm = TRUE)
            , total_child_beds_used_covid_confirmed = sum(total_pediatric_patients_hospitalized_confirmed_covid, na.rm = TRUE)
            , total_child_beds_used_covid = sum(total_pediatric_patients_hospitalized_confirmed_and_suspected_covid, na.rm = TRUE)
            , total_all_beds_used_covid = total_child_beds_used_covid + sum(total_adult_patients_hospitalized_confirmed_and_suspected_covid, na.rm = TRUE)
            , admission_child_covid_confirmed = sum(previous_day_admission_pediatric_covid_confirmed, na.rm = TRUE)
            , admission_child_covid = admission_child_covid_confirmed +
                                            sum(previous_day_admission_pediatric_covid_suspected, na.rm = TRUE)
            , admission_all_covid = admission_child_covid 
                                        + sum(previous_day_admission_adult_covid_suspected 
                                              + previous_day_admission_adult_covid_confirmed
                                              , na.rm = TRUE)
            , admission_confirmed_covid = admission_child_covid_confirmed 
                                        + sum(previous_day_admission_adult_covid_confirmed, na.rm = TRUE)
            , total_utilization = inpatient_all_beds_used/inpatient_all_beds
            , covid_deaths = sum(deaths_covid, na.rm = TRUE)
            , staffing_shortage_yes = sum(critical_staffing_shortage_today == 'true', na.rm = TRUE)
            , total_hosp_reporting = n()
            , .groups = 'keep')

# moving average of the variables above
state_ma <- state_d[order(state_d$state, state_d$collection_date),] %>%
  group_by(state) %>%
  mutate(
            total_all_beds = rollmean(total_all_beds, k=7, fill=NA)
            , inpatient_all_beds = rollmean(inpatient_all_beds, k=7, fill=NA)
            , icu_all_beds = rollmean(icu_all_beds, k=7, fill=NA)
            , total_child_beds = rollmean(total_child_beds, k=7, fill=NA)
            , inpatient_child_beds = rollmean(inpatient_child_beds, k=7, fill=NA)
            , icu_child_beds = rollmean(icu_child_beds, k=7, fill=NA)
            , total_all_beds_used_covid = rollmean(total_all_beds_used_covid, k=7, fill=NA)
            , total_child_beds_used_covid_confirmed = rollmean(total_child_beds_used_covid_confirmed, k=7, fill=NA)
            , total_child_beds_used_covid = rollmean(total_child_beds_used_covid, k=7, fill=NA)
            , admission_all_covid = rollmean(admission_all_covid, k=7, fill=NA)
            , admission_child_covid_confirmed = rollmean(admission_child_covid_confirmed, k=7, fill=NA)
            , admission_child_covid = rollmean(admission_child_covid, k=7, fill=NA)
            , total_utilization = rollmean(total_utilization, k=7, fill=NA)
            , covid_deaths = rollmean(covid_deaths, k=7, fill=NA)
            ) 

state <- state_ma %>%
  pivot_longer(cols = -c(state, collection_date)
               , names_pattern = "(...*)_(...*)_beds", names_to = c("bedtype", "scope")) %>%
  mutate(bedtype=ifelse(bedtype=="", "value", bedtype)) 
# %>%  pivot_wider(names_from = 'scope', values_from = 'value')
```

## State Level Weekly Average Data for the Rates
```{r}
daily_cases_by_state_rates <- daily_cases_by_state_rates %>%
  group_by(Province_State) %>%
  mutate(
    week_year = paste(week(date), year(date), sep=" ")   
    )

daily_cases_by_state_rates$state <- state.abb[match(daily_cases_by_state_rates$Province_State, state.name)]
daily_cases_by_state_rates <- daily_cases_by_state_rates %>%
  mutate(state = ifelse(Province_State == "District of Columbia", "DC", state))

daily_cases_by_state_rates <- rename(daily_cases_by_state_rates, collection_date = date)
daily_cases_by_state_rates <- daily_cases_by_state_rates[ , -1]

state_ma_rates <- state_d[order(state_d$state, state_d$collection_date),] %>%
  group_by(state) %>%
  mutate(
    week_year = paste(week(collection_date), year(collection_date), sep=" ")
            ) 

daily_cases_hosp_rates_full <- state_ma_rates %>%
  full_join(daily_cases_by_state_rates, by = c("state","collection_date"))

daily_cases_hosp_rates_full <- daily_cases_hosp_rates_full %>%
  group_by(state, week_year.y) %>%
  summarise(
    weekly_cases = mean(daily_cases, na.rm = TRUE)
    , weekly_all_covid_admissions = mean(admission_all_covid, na.rm = TRUE)
    , weekly_confirmed_covid_admissions = mean(admission_confirmed_covid, na.rm = TRUE)
    , weekly_covid_deaths = mean(covid_deaths, na.rm = TRUE)
    , date = first(collection_date)
    , lag_cases = lag(weekly_cases, default = first(weekly_cases), k=2)
    , lag_cases_death = lag(weekly_cases, default = first(weekly_cases), k=4)
    , .groups = 'keep'
            ) 
```

## Staffing Shortage Analysis (Slide 5)
```{r, fig.height=4, fig.width=6}
# Number of Hospitals Reporting lately
state_d %>%
  filter( state %in% target_states & collection_date > '2022-01-13') %>% 
  mutate(r = staffing_shortage_yes/total_hosp_reporting) %>%
  select(state, collection_date, r, total_hosp_reporting)

# Plot % of Hospitals per state reporting staffing shortages
state_d %>%
  # filter( state %in% target_states & total_hosp_reporting > 5 & collection_date > '2020-09-01') %>% 
  filter( state %in% target_states & collection_date > '2020-09-01' & collection_date <= date_most_hospitals_reported) %>% 
  mutate(r = staffing_shortage_yes/total_hosp_reporting) %>%
  ggplot(aes(x = collection_date, y = r)) + #, color = state
    facet_wrap(~ state, ncol = 2) + #scales = 'free_y' - For different axes
    geom_point(color = "grey55", size = 1) +
    geom_smooth(color = "blue", size = 1, span = 0.2) +
    labs(x = NULL, y= NULL, title = 'Percent of Hospitals Reporting Staffing Shortages') +
    # scale_x_date(date_breaks = "2 month", date_labels =  "%b-%d-%Y") +
    scale_x_date(breaks = seq.Date(as.Date('2020-11-01'), as.Date('2022-01-01'), "2 month"), date_labels =  "%b %d %Y") +
    scale_y_continuous(labels = scales::percent) +
    theme(axis.text.x = element_text(angle = 60, hjust=1))
```

## Join HHS data w/ other sources for remaining slides
```{r}
# Census data
state_ma_census <- state_ma %>%
  inner_join(df_population, by = "state")

# JHU Case data
daily_cases_hosp <- state_ma_census %>%
  inner_join(daily_cases_by_state, by = c("state","collection_date"))

# JHU Truth + CDC Ensemble Forecast
truth_hosp <- load_truth(
  truth_source = "HealthData",
  target_variable = "inc hosp",
  locations = target_statefips # "US"
)

truth_hosp_ma <- truth_hosp %>%
  filter(abbreviation %in% target_states) %>% # Filter for R1 states
  arrange(target_end_date) %>%
  ungroup() %>% group_by(abbreviation) %>% 
  mutate(admissions = zoo::rollmean(value, k = 7, align = "right", fill = NA))

df_plot <- bind_rows(
            truth_hosp_ma 
            , forecasts_hosp %>%
                filter(type == 'quantile') %>%
                pivot_wider(names_from = quantile, names_prefix = 'quant', values_from = value)
      ) %>%
  filter(target_end_date > "2021-11-01" #& !is.na(quant0.5)
         ) 


```

## COVID Case Incidence per state (Slide 3)
```{r, fig.height=4, fig.width=6}
daily_cases_hosp %>%
  filter(state %in% target_states & collection_date > '2020-09-15') %>% 
  mutate(r = (daily_cases/Population)*100000) %>%
  ggplot(aes(x = collection_date, y = r)) + #, color = state
    facet_wrap(~ state, ncol = 2) + #scales = 'free_y' - For different axes
    # geom_line(color = "blue", size = 1) +
    geom_point(color = "gray45", size = 1) +
    geom_smooth(color = "blue", size = 1, span = 0.02, se = FALSE) +
    labs(x = NULL, y = 'COVID-19 Cases', title = 'COVID-19 Daily Cases per 100K People, 7-day Moving Average') +
    scale_x_date(breaks = seq.Date(as.Date('2020-11-01'), as.Date('2022-01-01'), "2 month"), date_labels =  "%b %d %Y") +
    theme(axis.text.x = element_text(angle = 60, hjust=1))

# Pull out most recent values
daily_case_most_current <- daily_cases_hosp %>%
  filter(state %in% target_states & collection_date == date_case_count) %>% 
  mutate(r = (daily_cases/Population)*100000) %>%
  select(r)
```

## State-level Summary/Resource Prioritization (Slide 2)
```{r}
state_summary <- daily_cases_hosp %>%
  filter(state %in% target_states) %>%
  filter(!is.na(total_utilization)) %>%   
  group_by(state) #%>%
  #mutate(maxdate = max(collection_date, na.rm = TRUE)) 
#if (length(unique(state_summary$maxdate)) > 1){
  #state_summary$maxdate = date_most_hospitals_reported # min(unique(state_summary$maxdate))
#  state_summary$maxdate = '2022-01-26' # min(unique(state_summary$maxdate))
#}

# Will need to change date
# For 2/2, RI is not reporting so went back to 2/1
cols <- c('state', paste('Hospital Utilization as of', state_reported))
state_summary <- state_summary %>%
  filter(collection_date == state_reported) %>%
  select(state, total_utilization)
colnames(state_summary) <- cols
  
tmp <- daily_cases_hosp %>%
  filter(state %in% target_states) %>%
  mutate(inc100k = (daily_cases/Population)*100000) %>%
  filter(!is.na(inc100k)) %>%
  group_by(state) #%>%
  #mutate(maxdate = max(collection_date, na.rm = TRUE)) 
#if (length(unique(tmp$maxdate)) > 1){
#  tmp$maxdate = min(unique(tmp$maxdate))
#}
cols <- c('state', paste('New Cases per 100K as of', date_case_count))
tmp <- tmp %>%
  filter(collection_date == date_case_count) %>%
  select(state, inc100k)
colnames(tmp) <- cols

state_summary <- state_summary %>%
  inner_join(tmp, by = c('state'))

  
tmp <- state_ma_census %>%
  filter(state %in% target_states) %>%
  mutate(r = (admission_all_covid/Population)*100000) %>%
  filter(!is.na(r)) %>%
  group_by(state) #%>%
  #mutate(maxdate = max(collection_date, na.rm = TRUE)) 
#if (length(unique(tmp$maxdate)) > 1){
#  tmp$maxdate = min(unique(tmp$maxdate))
#}
cols <- c('state', paste('Hosp. Admissions per 100K as of', date_most_hospitals_reported))
tmp <- tmp %>%
  filter(collection_date == date_most_hospitals_reported) %>%
  select(state, r)
colnames(tmp) <- cols

state_summary <- state_summary %>%
  inner_join(tmp, by = c('state'))

tmp <- state_d %>%
  filter( state %in% target_states) %>%
  mutate(r = staffing_shortage_yes/total_hosp_reporting) %>%
  filter(total_hosp_reporting > 5 & !is.na(r)) %>%
  group_by(state) %>%
  filter((state != 'MA' & collection_date == date_most_hospitals_reported)
         | (state == 'MA'& collection_date == date_most_hospitals_reported)) %>%
  select(state, r, collection_date)
colnames(tmp) <- c('state', 'Staffing Shortage', 'Shortage as of')

state_summary <- state_summary %>%
  inner_join(tmp, by = c('state')) #%>%
  #arrange(match(state, c("RI", "MA", "CT", "NH", "VT", "ME")))

state_summary %>%
  kableExtra::kbl() %>% kableExtra::kable_styling()
```


## CDC Ensemble Forecasts per State (Slide 4)
```{r, fig.height=4,fig.width=6}
plotFcst <- function(tit, df_plot, ylims){
  brks <- sort(c(seq.Date(as.Date('2020-11-01'), as.Date('2022-02-01'), "1 month"),
                seq.Date(as.Date('2020-11-15'), as.Date('2022-02-01'), "1 month")))
ggplot(df_plot, aes(x = target_end_date)) +
    facet_wrap(~ abbreviation, ncol = 2, scales = 'free_y') + # - For different axes
    geom_line(aes(y = admissions, color = "Reported (with 7-d mov. average"), size = 1) +
    geom_point(aes(y = value, color = "Reported (with 7-d mov. average"), size = 1.5) +
    # geom_smooth(aes(y = value), color = "blue", size = 1, span = 0.3, se = FALSE) +
    scale_color_manual(name = NULL, values = c("blue")) +
    geom_ribbon(aes(ymin = quant0.025, ymax = quant0.975, fill = "95% Prediction Interval"), size = 1) +
    geom_ribbon(aes(ymin = quant0.25, ymax = quant0.75, fill = "50% Prediction Interval"), size = 1) +
    scale_fill_manual(name = NULL, values = c("salmon3", "salmon1", "orangered4")) +
    geom_point(aes(y = quant0.5), color = "orangered4", size = 1) +
    geom_line(aes(y = quant0.5, fill = "Ensemble"), size = 1) +
    labs(x = NULL, y = NULL, title = tit) +
    expand_limits(y = ylims) +
    scale_x_date(breaks = brks, date_labels =  "%b %d %Y") +
    guides(color = guide_legend(order=1),
         fill = guide_legend(order=2)) +
    theme(axis.text.x = element_text(angle = 60, hjust=1))
}

plotFcst('New Hospital Admissions per day', df_plot, c(0, 100))
```

## COVID Deaths per State (Slide 8)
```{r, fig.height=4,fig.width=6}
# Load JHU Truth on R1 deaths
truth_death <- load_truth(
  truth_source = "JHU",
  target_variable = "inc death",
  locations = target_statefips
)

truth_death %>%
  filter(target_end_date > '2020-09-30') %>%
  mutate(r = value/population * 1e5) %>%
ggplot(aes(x = target_end_date, y = r)) +
    facet_wrap(~ abbreviation, ncol = 2) + # , scales = 'free_y' - For different axes
    geom_point(color = "grey45", size = 1) +
    geom_smooth(color = "blue", size = 1, span = 0.1, se = FALSE) +
    labs(x = NULL, y = NULL, title = 'COVID-19 Weekly Deaths per 100K people') +
    # expand_limits(y = ylims) +
    scale_x_date(breaks = seq.Date(as.Date('2020-11-01'), as.Date('2022-01-01'), "2 month"), date_labels =  "%b %d %Y") +
    theme(axis.text.x = element_text(angle = 60, hjust=1))
```
```{r}
truth_death %>%
  filter(target_end_date > '2022-01-10') %>%
  mutate(r = value/population * 1e5) %>%
  select(target_end_date, location_name, r, value)
```
## Get National valus for cases and deaths
```{r}
# National COVID Deaths
truth_death_US <- load_truth(
  truth_source = "JHU",
  target_variable = "inc death",
  locations = "US"
)
  
truth_death_US  %>%
  tail(n = 1) %>%
  mutate(per100K = value / population * 1e5) %>%
  select(model, location, target_end_date, value, population, per100K)

# National Incidence Data
truth_incidence_US <- load_truth(
  truth_source = "JHU",
  target_variable = "inc case",
  locations = "US"
)

truth_incidence_US %>%
  mutate(incday100K = value / population * 1e5 / 7) %>%
  filter(target_end_date == max(target_end_date)) %>%
  select(location, target_end_date, value, incday100K)
```


# HRR-level Analysis

## Aggregate by HRR
```{r}
hrr_d <- df_raw %>%
  group_by(hrr, collection_date) %>%
  summarise(
            total_all_beds = sum(total_beds, na.rm = TRUE)
            , inpatient_all_beds = sum(inpatient_beds, na.rm = TRUE)
            , icu_all_beds = sum(total_icu_beds, na.rm = TRUE)
            , total_child_beds = total_all_beds - sum(all_adult_hospital_beds, na.rm = TRUE)
            , inpatient_child_beds = inpatient_all_beds - sum(all_adult_hospital_inpatient_beds, na.rm = TRUE)
            , icu_child_beds = icu_all_beds - sum(total_staffed_adult_icu_beds, na.rm = TRUE)
            , inpatient_all_beds_used = sum(inpatient_beds_used, na.rm = TRUE)
            , inpatient_child_beds_used = inpatient_all_beds_used - sum(all_adult_hospital_inpatient_bed_occupied, na.rm = TRUE)
            , icu_all_beds_used = sum(icu_beds_used, na.rm = TRUE)
            , icu_child_beds_used = icu_all_beds_used - sum(staffed_adult_icu_bed_occupancy, na.rm = TRUE)
            , total_child_beds_used_covid_confirmed = sum(total_pediatric_patients_hospitalized_confirmed_covid, na.rm = TRUE)
            , total_child_beds_used_covid = sum(total_pediatric_patients_hospitalized_confirmed_and_suspected_covid, na.rm = TRUE)
            , total_all_beds_used_covid = total_child_beds_used_covid + sum(total_adult_patients_hospitalized_confirmed_and_suspected_covid, na.rm = TRUE)
            , admission_child_covid_confirmed = sum(previous_day_admission_pediatric_covid_confirmed, na.rm = TRUE)
            , admission_child_covid = admission_child_covid_confirmed +
                                            sum(previous_day_admission_pediatric_covid_suspected, na.rm = TRUE)
            , admission_all_covid = admission_child_covid 
                                        + sum(previous_day_admission_adult_covid_suspected 
                                              + previous_day_admission_adult_covid_confirmed
                                              , na.rm = TRUE)
            , admission_confirmed_covid = admission_child_covid_confirmed 
                                        + sum(previous_day_admission_adult_covid_confirmed, na.rm = TRUE)
            , non_child_covid_utilization = ((icu_all_beds_used + inpatient_all_beds_used) - total_child_beds_used_covid)/total_all_beds
            , non_covid_utilization = (inpatient_all_beds_used - total_all_beds_used_covid)/inpatient_all_beds
            , covid_utilization = total_all_beds_used_covid/inpatient_all_beds
            , util_ratio = covid_utilization/non_covid_utilization
            # , total_utilization = (inpatient_all_beds_used)/total_all_beds
            , total_utilization = inpatient_all_beds_used/inpatient_all_beds
            , covid_ratio = covid_utilization/total_utilization
            , icu_utilization = icu_all_beds_used/icu_all_beds
            , general_utilization = (inpatient_all_beds_used-icu_all_beds_used)/(inpatient_all_beds-icu_all_beds)
            , covid_deaths = sum(deaths_covid, na.rm = TRUE)
            , .groups = 'keep')

# moving average of the variables above
hrr_ma <- hrr_d[order(hrr_d$hrr, hrr_d$collection_date),] %>%
  group_by(hrr) %>%
  mutate(
            total_all_beds = rollmean(total_all_beds, k=7, fill=NA)
            , inpatient_all_beds = rollmean(inpatient_all_beds, k=7, fill=NA)
            , icu_all_beds = rollmean(icu_all_beds, k=7, fill=NA)
            , total_child_beds = rollmean(total_child_beds, k=7, fill=NA)
            , inpatient_child_beds = rollmean(inpatient_child_beds, k=7, fill=NA)
            , icu_child_beds = rollmean(icu_child_beds, k=7, fill=NA)
            , total_all_beds_used_covid = rollmean(total_all_beds_used_covid, k=7, fill=NA)
            , total_child_beds_used_covid_confirmed = rollmean(total_child_beds_used_covid_confirmed, k=7, fill=NA)
            , total_child_beds_used_covid = rollmean(total_child_beds_used_covid, k=7, fill=NA)
            , admission_all_covid = rollmean(admission_all_covid, k=7, fill=NA)
            , admission_child_covid_confirmed = rollmean(admission_child_covid_confirmed, k=7, fill=NA)
            , admission_child_covid = rollmean(admission_child_covid, k=7, fill=NA)
            , non_child_covid_utilization = rollmean(non_child_covid_utilization, k=7, fill=NA)
            , total_utilization = rollmean(total_utilization, k=7, fill=NA)
            , icu_utilization = rollmean(icu_utilization, k=7, fill=NA)
            , general_utilization = rollmean(general_utilization, k=7, fill=NA)
            , covid_deaths = rollmean(covid_deaths, k=7, fill=NA)
            ) 
```

## Divide into Region 1 HRRs
```{r}
hrr_d_NE <- hrr_d %>%
  filter( hrr %in% c(109,110,111,221,222,227,230,231,281,282,295,364,424) & collection_date > '2020-09-30') 
hrr_d_NE$hrr <- factor(hrr_d_NE$hrr, levels = c(109,110,111,221,222,227,230,231,281,282,364,424,295))
hrr_d_NE <- hrr_d_NE%>%
  mutate(hrr=recode(hrr, 
                        '109' = 'CT - Bridgeport',
                        '110'	= 'CT - Hartford',
                        '111'	= 'CT - New Haven',
                        '221'	= 'ME - Bangor',
                        '222'	= 'ME - Portland',
                        '227'	= 'MA - Boston',
                        '230'	= 'MA - Springfield',
                        '231'	= 'MA - Worcester',
                        '281'	= 'NH - Lebanon',
                        '282'	= 'NH - Manchester',
                        '295'	= 'NY - Albany',
                        '364'	= 'RI - Providence',
                        '424'	= 'VT - Burlington'))
```

## Hospital Utilization Rates by HRR (Slide 6)
```{r, fig.height=4,fig.width=6}
hrr_d_NE %>%
  filter(collection_date <= date_most_hospitals_reported) %>%
  ggplot(aes(x = collection_date, y = total_utilization)) + #, color = state
    facet_wrap(~ hrr, ncol = 3) + #scales = 'free_y' - For different axes
    geom_hline(yintercept = 0.8, color = 'orange', linetype='dashed') +
    geom_line(color = "blue", size = 1) +
    # geom_point(color = "grey", size = 0.5) +
    # geom_smooth(color = "blue", size = 1, span = 0.01) +
    labs(x = NULL, y = NULL, title = 'COVID-19 Total Hospital Utilization') +
    # scale_x_date(date_breaks = "2 month", date_labels =  "%b %d %Y") +
    scale_x_date(breaks = seq.Date(as.Date('2020-11-01'), as.Date('2022-01-01'), "2 month"), date_labels =  "%b %d %Y") +
    scale_y_continuous(labels=scales::percent, limits = c(0, 1)) +
    theme( text = element_text(size=12)) +
    theme(axis.text.x = element_text(angle = 60, hjust=1))
```
## Pull latest HRR Utilization rates
```{r}
hrr_last_util <- hrr_d_NE %>%
  filter(collection_date == date_most_hospitals_reported) %>%
  select(hrr,collection_date,total_utilization)

fname <- paste0(date_most_hospitals_reported ,'_HRR_last_util.xlsx')

openxlsx::write.xlsx( 
              hrr_last_util
              , fname )
```


## COVID vs. Non-COVID utilization
```{r}
hrr_d_NE %>%
  ggplot() + # color = state
    facet_wrap(~ hrr, ncol = 3) + #scales = 'free_y' - For different axes
    geom_line(aes(x = collection_date, y = total_utilization, color="Total Utilization")) + 
    geom_line(aes(x = collection_date, y = non_covid_utilization, color = "Non-COVID-19 Utilization")) +
    geom_ribbon(aes(ymin = non_covid_utilization, ymax = total_utilization, x=collection_date), fill="blue") + 
    labs(x = NULL, y = NULL, title = 'COVID-19 Hospital Utilization') +
    scale_x_date(breaks = seq.Date(as.Date('2020-11-01'), as.Date('2022-01-01'), "2 month"), date_labels =  "%b %d %Y") +
    scale_y_continuous(labels=scales::percent) +
    theme( text = element_text(size=12)) +
    theme(axis.text.x = element_text(angle = 60, hjust=1))
```

## Plot COVID utilization
```{r, fig.height=4,fig.width=6}
hrr_d_NE %>%
  filter(collection_date <= date_most_hospitals_reported) %>%
  ggplot(aes(x = collection_date, y = covid_utilization)) + #, color = state
    facet_wrap(~ hrr, ncol = 3) + #scales = 'free_y' - For different axes
    #geom_hline(yintercept = 0.8, color = 'orange', linetype='dashed') +
    geom_line(color = "blue", size = 1) +
    # geom_point(color = "grey", size = 0.5) +
    # geom_smooth(color = "blue", size = 1, span = 0.01) +
    labs(x = NULL, y = NULL, title = 'COVID Utilization per HRR') +
    # scale_x_date(date_breaks = "2 month", date_labels =  "%b %d %Y") +
    scale_x_date(breaks = seq.Date(as.Date('2020-11-01'), as.Date('2022-01-01'), "2 month"), date_labels =  "%b %d %Y") +
    scale_y_continuous(labels=scales::percent) +
    theme( text = element_text(size=12)) +
    theme(axis.text.x = element_text(angle = 60, hjust=1))
```



## Plot Ratio of COVID to non-COVID utilization
```{r}
hrr_d_NE %>%
  filter(collection_date <= date_most_hospitals_reported) %>%
  ggplot(aes(x = collection_date, y = util_ratio)) + #, color = state
    facet_wrap(~ hrr, ncol = 3) + #scales = 'free_y' - For different axes
    #geom_hline(yintercept = 0.8, color = 'orange', linetype='dashed') +
    geom_line(color = "blue", size = 1) +
    # geom_point(color = "grey", size = 0.5) +
    # geom_smooth(color = "blue", size = 1, span = 0.01) +
    labs(x = NULL, y = NULL, title = 'Ratio of COVID to non-COVID Hospital Utilization') +
    # scale_x_date(date_breaks = "2 month", date_labels =  "%b %d %Y") +
    scale_x_date(breaks = seq.Date(as.Date('2020-11-01'), as.Date('2022-01-01'), "2 month"), date_labels =  "%b %d %Y") +
    #scale_y_continuous(labels=scales::percent, limits = c(0, 1)) +
    theme( text = element_text(size=12)) +
    theme(axis.text.x = element_text(angle = 60, hjust=1))

hrr_d_NE %>%
  filter(collection_date <= date_most_hospitals_reported) %>%
  ggplot(aes(x = collection_date, y = covid_ratio)) + #, color = state
    facet_wrap(~ hrr, ncol = 3) + #scales = 'free_y' - For different axes
    #geom_hline(yintercept = 0.8, color = 'orange', linetype='dashed') +
    geom_line(color = "blue", size = 1) +
    # geom_point(color = "grey", size = 0.5) +
    # geom_smooth(color = "blue", size = 1, span = 0.01) +
    labs(x = NULL, y = NULL, title = 'COVID as a percentage of Total Hospital Utilization') +
    # scale_x_date(date_breaks = "2 month", date_labels =  "%b %d %Y") +
    scale_x_date(breaks = seq.Date(as.Date('2020-11-01'), as.Date('2022-01-01'), "2 month"), date_labels =  "%b %d %Y") +
    scale_y_continuous(labels=scales::percent) +
    theme( text = element_text(size=12)) +
    theme(axis.text.x = element_text(angle = 60, hjust=1))
```


# Investigate Individual States - RI

## Filter data by target state
```{r}
RI_raw <- df_raw %>%
  filter(state %in% c("RI"))
```

## Filter by facility in target states
```{r}
RI_d_h <- RI_raw %>%
  group_by(hospital_name, collection_date, state) %>%
  summarise(
            total_all_beds = sum(total_beds, na.rm = TRUE)
            , inpatient_all_beds = sum(inpatient_beds, na.rm = TRUE)
            , icu_all_beds = sum(total_icu_beds, na.rm = TRUE)
            , total_child_beds = total_all_beds - sum(all_adult_hospital_beds, na.rm = TRUE)
            , inpatient_child_beds = inpatient_all_beds - sum(all_adult_hospital_inpatient_beds, na.rm = TRUE)
            , icu_child_beds = icu_all_beds - sum(total_staffed_adult_icu_beds, na.rm = TRUE)
            , inpatient_all_beds_used = sum(inpatient_beds_used, na.rm = TRUE)
            , inpatient_child_beds_used = inpatient_all_beds_used - sum(all_adult_hospital_inpatient_bed_occupied, na.rm = TRUE)
            , icu_all_beds_used = sum(icu_beds_used, na.rm = TRUE)
            , icu_child_beds_used = icu_all_beds_used - sum(staffed_adult_icu_bed_occupancy, na.rm = TRUE)
            , total_child_beds_used_covid_confirmed = sum(total_pediatric_patients_hospitalized_confirmed_covid, na.rm = TRUE)
            , total_child_beds_used_covid = sum(total_pediatric_patients_hospitalized_confirmed_and_suspected_covid, na.rm = TRUE)
            , total_all_beds_used_covid = total_child_beds_used_covid + sum(total_adult_patients_hospitalized_confirmed_and_suspected_covid, na.rm = TRUE)
            , admission_child_covid_confirmed = sum(previous_day_admission_pediatric_covid_confirmed, na.rm = TRUE)
            , admission_child_covid = admission_child_covid_confirmed +
                                            sum(previous_day_admission_pediatric_covid_suspected, na.rm = TRUE)
            , admission_all_covid = admission_child_covid 
                                        + sum(previous_day_admission_adult_covid_suspected 
                                              + previous_day_admission_adult_covid_confirmed
                                              , na.rm = TRUE)
            , non_child_covid_utilization = (inpatient_all_beds_used - total_child_beds_used_covid)/total_all_beds
            , non_covid_utilization = (inpatient_all_beds_used - total_all_beds_used_covid)/inpatient_all_beds
            , total_utilization = (inpatient_all_beds_used)/inpatient_all_beds
            , covid_deaths = sum(deaths_covid, na.rm = TRUE)
            , staffing_shortage_yes = sum(critical_staffing_shortage_today == 'yes', na.rm = TRUE)
            , .groups = 'keep')

# moving average of the variables above
RI_ma_h <- RI_d_h[order(RI_d_h$hospital_name, RI_d_h$collection_date, RI_d_h$state),] %>%
  group_by(hospital_name) %>%
  mutate(
            total_all_beds = rollmean(total_all_beds, k=7, fill=NA)
            , inpatient_all_beds = rollmean(inpatient_all_beds, k=7, fill=NA)
            , icu_all_beds = rollmean(icu_all_beds, k=7, fill=NA)
            , total_child_beds = rollmean(total_child_beds, k=7, fill=NA)
            , inpatient_child_beds = rollmean(inpatient_child_beds, k=7, fill=NA)
            , icu_child_beds = rollmean(icu_child_beds, k=7, fill=NA)
            , total_all_beds_used_covid = rollmean(total_all_beds_used_covid, k=7, fill=NA)
            , total_child_beds_used_covid_confirmed = rollmean(total_child_beds_used_covid_confirmed, k=7, fill=NA)
            , total_child_beds_used_covid = rollmean(total_child_beds_used_covid, k=7, fill=NA)
            , admission_all_covid = rollmean(admission_all_covid, k=7, fill=NA)
            , admission_child_covid_confirmed = rollmean(admission_child_covid_confirmed, k=7, fill=NA)
            , admission_child_covid = rollmean(admission_child_covid, k=7, fill=NA)
            , non_child_covid_utilization = rollmean(non_child_covid_utilization, k=7, fill=NA)
            , non_covid_utilization = rollmean(non_covid_utilization, k=7, fill=NA)
            , total_utilization = rollmean(total_utilization, k=7, fill=NA)
            , covid_deaths = rollmean(covid_deaths, k=7, fill=NA)
            ) 
```

## Target State Hosptial utilization
```{r}
ri_hosp <- c("KENT COUNTY MEMORIAL HOSPITAL", "LANDMARK MEDICAL CENTER", "RHODE ISLAND HOSPITAL", "THE MIRIAM HOSPITAL","OUR LADY OF FATIMA HOSPITAL", "NEWPORT HOSPITAL", "WESTERLY HOSPITAL", "ROGER WILLIAMS MEDICAL CENTER", "SOUTH COUNTY HOSPITAL INC")

hospital_df <- RI_d_h %>%
  filter(collection_date == date_most_hospitals_reported & hospital_name %in% ri_hosp) %>%
  mutate(inp_percent_utilization = inpatient_all_beds_used/inpatient_all_beds) %>%
  mutate(icu_percent_utilization = icu_all_beds_used/icu_all_beds) %>%
  mutate(total_percent_utilization = (inpatient_all_beds_used+icu_all_beds_used)/(inpatient_all_beds+icu_all_beds)) %>%
  select(inp_percent_utilization,icu_percent_utilization,total_percent_utilization,staffing_shortage_yes)
```


```{r}
hospital_df <- RI_d_h %>%
  filter(collection_date == date_most_hospitals_reported)
```


## Write to Excel file (RI Slide 7)
```{r}
fname1 <- paste0('2022-02-02_RI_hosp_util.xlsx')

openxlsx::write.xlsx( 
              hospital_df
              , fname1 )
```

## Plot Utilization for each RI facility
```{r, fig.height=4, fig.width=6}
state_d_h %>%
  filter(hospital_name %in% ri_hosp) %>%
  filter(collection_date <= date_most_hospitals_reported) %>%
  ggplot(aes(x = collection_date, y = total_utilization)) + #, color = state
    facet_wrap(~ hospital_name, ncol = 3) + #scales = 'free_y' - For different axes
    geom_hline(yintercept = 0.8, color = 'orange', linetype='dashed') +
    geom_line(color = "blue", size = 1) +
    # geom_point(color = "grey", size = 0.5) +
    # geom_smooth(color = "blue", size = 1, span = 0.01) +
    labs(x = NULL, y = NULL, title = 'COVID-19 Total Hospital Utilization') +
    # scale_x_date(date_breaks = "2 month", date_labels =  "%b %d %Y") +
    scale_x_date(breaks = seq.Date(as.Date('2020-11-01'), as.Date('2022-01-01'), "2 month"), date_labels =  "%b %d %Y") +
    scale_y_continuous(labels=scales::percent, limits = c(0, 1)) +
    theme( text = element_text(size=12))
```

