---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(zoo)
library(readxl)

theme_set(theme_bw() + theme( legend.position="bottom" ) +
             theme( text = element_text(size=16, face="bold")) +
            theme(axis.text.x=element_text(angle=60, hjust=1)))

# R1states <- c('CT', 'MA', 'ME', 'NH', 'RI', 'VT') # FEMA Region 1
# R2states <- c('NY', 'NJ', 'PR')
# R3states <- c('DE', 'DC', 'MD', 'PA', 'VA', 'WV')
# R4states <- c('AL', 'FL', 'GA', 'KY', 'MS', 'NC', 'SC', 'TN') # Region 4
# R5states <- c('IL', 'IN', 'MI', 'MN', 'OH', 'WI')
# R6states <- c('TX', 'AR', 'LA', 'NM', 'OK')
# R7states <- c('IA', 'KS', 'MO', 'NE', '')
# R8states <- c('CO', 'MT', 'ND', 'SD', 'UT', 'WY')
# R9states <- c('AZ', 'CA', 'HI', 'NV') # Some Islands Missing Abriviations
# R10states <- c('AK', 'ID', 'OR', 'WA')

target_states <- c('DC', 'MS', 'NC', 'AK', 'NH', 'IL') # c('FL', 'MS', 'TX', 'CA') # Non-New England States to Compare
```

load data from Census Data: https://datacenter.kidscount.org/data/tables/101-child-population-by-age-group#detailed/2/2-53/false/1729,37/62,63,64,6,4693/419
https://datacenter.kidscount.org/rawdata.axd?ind=101&loc=1
```{r}
df_child_counts <- read_excel('data/Child population by age group.xlsx') 
```

Clean up/Select Data
```{r}
df_child_counts$state <- state.abb[match(df_child_counts$Location,state.name)]
df_child_counts <- df_child_counts %>%
  mutate(state = ifelse(Location == "District of Columbia", "DC", state))

df_child_counts <- df_child_counts %>%
  mutate(region = case_when(
    # state %in% R1states ~ "Region 1",
    # state %in% R2states ~ "Region 2",
    # state %in% R3states ~ "Region 3",
    # state %in% R4states ~ "Region 4",
    # state %in% R5states ~ "Region 5",
    # state %in% R6states ~ "Region 6",
    # state %in% R7states ~ "Region 7",
    # state %in% R8states ~ "Region 8",
    # state %in% R9states ~ "Region 9",
    # state %in% R10states ~ "Region 10",
    state %in% target_states ~ "Selected States"
    ))

df_child_counts$Data <- as.numeric(df_child_counts$Data) 

df_2019child_counts <- df_child_counts %>%
                        filter(TimeFrame == '2019' & (LocationType == 'State' | Location == "District of Columbia")
                              & DataFormat == 'Number' & `Age group` == "Total less than 18") %>%
                        select(-"LocationType", -"DataFormat", -"TimeFrame")

df_2019child_counts_r <- df_2019child_counts %>%
  group_by(region) %>%
  summarise(
    child_population = sum(Data, na.rm = TRUE)
    , .groups = 'keep'
  )
```

load data from HHS Protect obtained by Jeff Cegan
```{r}
df_raw <- read.csv('data/[Unified] Hospital Analytic.csv') 
```

```{r}
facility <- df_raw %>%
  mutate(collection_date = as.Date(collection_date))
```

```{r}
summary(facility %>%
  select(hospital_pk, collection_date, total_beds, inpatient_beds, total_icu_beds)
  )
```

Aggregate by state
```{r}
state_d <- facility %>%
  group_by(state, collection_date) %>%
  summarise(
            total_all_beds = sum(total_beds, na.rm = TRUE)
            , inpatient_all_beds = sum(inpatient_beds, na.rm = TRUE)
            , icu_all_beds = sum(total_icu_beds, na.rm = TRUE)
            , total_child_beds = total_all_beds - sum(all_adult_hospital_beds, na.rm = TRUE)
            , inpatient_child_beds = inpatient_all_beds - sum(all_adult_hospital_inpatient_beds, na.rm = TRUE)
            , icu_child_beds = icu_all_beds - sum(total_staffed_adult_icu_beds, na.rm = TRUE)
            , inpatient_all_beds_used = sum(inpatient_beds_used, na.rm = TRUE)
            , inpatient_child_beds_used = inpatient_all_beds_used - sum(all_adult_hospital_inpatient_bed_occupied, na.rm = TRUE)
            , icu_all_beds_used = sum(icu_beds_used, na.rm = TRUE)
            , icu_child_beds_used = icu_all_beds_used - sum(staffed_adult_icu_bed_occupancy, na.rm = TRUE)
            , total_child_beds_used_covid_confirmed = sum(total_pediatric_patients_hospitalized_confirmed_covid, na.rm = TRUE)
            , total_child_beds_used_covid = sum(total_pediatric_patients_hospitalized_confirmed_and_suspected_covid, na.rm = TRUE)
            , total_all_beds_used_covid = total_child_beds_used_covid + sum(total_adult_patients_hospitalized_confirmed_and_suspected_covid, na.rm = TRUE)
            , admission_child_covid_confirmed = sum(previous_day_admission_pediatric_covid_confirmed, na.rm = TRUE)
            , admission_child_covid = admission_child_covid_confirmed +
                                            sum(previous_day_admission_pediatric_covid_suspected, na.rm = TRUE)
            , admission_all_covid = admission_child_covid 
                                        + sum(previous_day_admission_adult_covid_suspected 
                                              + previous_day_admission_adult_covid_confirmed
                                              , na.rm = TRUE)
            , non_child_covid_utilization = ((inpatient_all_beds_used) - total_child_beds_used_covid)/total_all_beds
            # , total_utilization = (inpatient_all_beds_used)/total_all_beds
            , total_utilization = (inpatient_all_beds_used)/inpatient_all_beds
            , .groups = 'keep')

# moving average of the variables above
state_ma <- state_d[order(state_d$state, state_d$collection_date),] %>%
  group_by(state) %>%
  mutate(
            total_all_beds = rollmean(total_all_beds, k=7, fill=NA)
            , inpatient_all_beds = rollmean(inpatient_all_beds, k=7, fill=NA)
            , icu_all_beds = rollmean(icu_all_beds, k=7, fill=NA)
            , total_child_beds = rollmean(total_child_beds, k=7, fill=NA)
            , inpatient_child_beds = rollmean(inpatient_child_beds, k=7, fill=NA)
            , icu_child_beds = rollmean(icu_child_beds, k=7, fill=NA)
            , total_all_beds_used_covid = rollmean(total_all_beds_used_covid, k=7, fill=NA)
            , total_child_beds_used_covid_confirmed = rollmean(total_child_beds_used_covid_confirmed, k=7, fill=NA)
            , total_child_beds_used_covid = rollmean(total_child_beds_used_covid, k=7, fill=NA)
            , admission_all_covid = rollmean(admission_all_covid, k=7, fill=NA)
            , admission_child_covid_confirmed = rollmean(admission_child_covid_confirmed, k=7, fill=NA)
            , admission_child_covid = rollmean(admission_child_covid, k=7, fill=NA)
            , non_child_covid_utilization = rollmean(non_child_covid_utilization, k=7, fill=NA)
            , total_utilization = rollmean(total_utilization, k=7, fill=NA)
            ) 

state <- state_ma %>%
  pivot_longer(cols = -c(state, collection_date)
               , names_pattern = "(...*)_(...*)_beds", names_to = c("bedtype", "scope")) %>%
  mutate(bedtype=ifelse(bedtype=="", "value", bedtype)) 
# %>%  pivot_wider(names_from = 'scope', values_from = 'value')
```

Aggregate by Region
```{r}
region_d <- facility %>%
  group_by(fema_region, collection_date) %>%
  summarise(
            total_all_beds = sum(total_beds, na.rm = TRUE)
            , inpatient_all_beds = sum(inpatient_beds, na.rm = TRUE)
            , icu_all_beds = sum(total_icu_beds, na.rm = TRUE)
            , total_child_beds = total_all_beds - sum(all_adult_hospital_beds, na.rm = TRUE)
            , inpatient_child_beds = inpatient_all_beds - sum(all_adult_hospital_inpatient_beds, na.rm = TRUE)
            , icu_child_beds = icu_all_beds - sum(total_staffed_adult_icu_beds, na.rm = TRUE)
            , inpatient_all_beds_used = sum(inpatient_beds_used, na.rm = TRUE)
            , inpatient_child_beds_used = inpatient_all_beds_used - sum(all_adult_hospital_inpatient_bed_occupied, na.rm = TRUE)
            , icu_all_beds_used = sum(icu_beds_used, na.rm = TRUE)
            , icu_child_beds_used = icu_all_beds_used - sum(staffed_adult_icu_bed_occupancy, na.rm = TRUE)
            , total_child_beds_used_covid_confirmed = sum(total_pediatric_patients_hospitalized_confirmed_covid, na.rm = TRUE)
            , total_child_beds_used_covid = sum(total_pediatric_patients_hospitalized_confirmed_and_suspected_covid, na.rm = TRUE)
            , total_all_beds_used_covid = total_child_beds_used_covid + sum(total_adult_patients_hospitalized_confirmed_and_suspected_covid, na.rm = TRUE)
            , admission_child_covid_confirmed = sum(previous_day_admission_pediatric_covid_confirmed, na.rm = TRUE)
            , admission_child_covid = admission_child_covid_confirmed +
                                            sum(previous_day_admission_pediatric_covid_suspected, na.rm = TRUE)
            , admission_all_covid = admission_child_covid 
                                        + sum(previous_day_admission_adult_covid_suspected 
                                              + previous_day_admission_adult_covid_confirmed
                                              , na.rm = TRUE)
            , non_child_covid_utilization = ((inpatient_all_beds_used) - total_child_beds_used_covid)/total_all_beds
            , total_utilization = (inpatient_all_beds_used)/total_all_beds
            , .groups = 'keep')

# moving average of the variables above
region_ma <- region_d[order(region_d$fema_region, region_d$collection_date),] %>%
  group_by(fema_region) %>%
  mutate(
            total_all_beds = rollmean(total_all_beds, k=7, fill=NA)
            , inpatient_all_beds = rollmean(inpatient_all_beds, k=7, fill=NA)
            , icu_all_beds = rollmean(icu_all_beds, k=7, fill=NA)
            , total_child_beds = rollmean(total_child_beds, k=7, fill=NA)
            , inpatient_child_beds = rollmean(inpatient_child_beds, k=7, fill=NA)
            , icu_child_beds = rollmean(icu_child_beds, k=7, fill=NA)
            , total_all_beds_used_covid = rollmean(total_all_beds_used_covid, k=7, fill=NA)
            , total_child_beds_used_covid_confirmed = rollmean(total_child_beds_used_covid_confirmed, k=7, fill=NA)
            , total_child_beds_used_covid = rollmean(total_child_beds_used_covid, k=7, fill=NA)
            , admission_all_covid = rollmean(admission_all_covid, k=7, fill=NA)
            , admission_child_covid_confirmed = rollmean(admission_child_covid_confirmed, k=7, fill=NA)
            , admission_child_covid = rollmean(admission_child_covid, k=7, fill=NA)
            , non_child_covid_utilization = rollmean(non_child_covid_utilization, k=7, fill=NA)
            , total_utilization = rollmean(total_utilization, k=7, fill=NA)
            ) 

region <- region_ma %>%
  pivot_longer(cols = -c(fema_region, collection_date)
               , names_pattern = "(...*)_(...*)_beds", names_to = c("bedtype", "scope")) %>%
  mutate(bedtype=ifelse(bedtype=="", "value", bedtype)) 
# %>%  pivot_wider(names_from = 'scope', values_from = 'value')
```

Children's Hospitals
```{r}
state_d_ch <- facility %>%
  filter(hospital_subtype == 'Childrens Hospitals') %>%
  group_by(hospital_name, collection_date) %>%
  summarise(
            inpatient_all_beds = sum(inpatient_beds, na.rm = TRUE)
            , icu_all_beds = sum(total_icu_beds, na.rm = TRUE)
            , total_all_beds = inpatient_all_beds
            , total_child_beds = total_all_beds - sum(all_adult_hospital_beds, na.rm = TRUE)
            , inpatient_child_beds = inpatient_all_beds - sum(all_adult_hospital_inpatient_beds, na.rm = TRUE)
            , icu_child_beds = icu_all_beds - sum(total_staffed_adult_icu_beds, na.rm = TRUE)
            , inpatient_all_beds_used = sum(inpatient_beds_used, na.rm = TRUE)
            , inpatient_child_beds_used = inpatient_all_beds_used - sum(all_adult_hospital_inpatient_bed_occupied, na.rm = TRUE)
            , icu_all_beds_used = sum(icu_beds_used, na.rm = TRUE)
            , icu_child_beds_used = icu_all_beds_used - sum(staffed_adult_icu_bed_occupancy, na.rm = TRUE)
            , total_child_beds_used_covid_confirmed = sum(total_pediatric_patients_hospitalized_confirmed_covid, na.rm = TRUE)
            , total_child_beds_used_covid = sum(total_pediatric_patients_hospitalized_confirmed_and_suspected_covid, na.rm = TRUE)
            , total_all_beds_used_covid = total_child_beds_used_covid + sum(total_adult_patients_hospitalized_confirmed_and_suspected_covid, na.rm = TRUE)
            , admission_child_covid_confirmed = sum(previous_day_admission_pediatric_covid_confirmed, na.rm = TRUE)
            , admission_child_covid = admission_child_covid_confirmed +
                                            sum(previous_day_admission_pediatric_covid_suspected, na.rm = TRUE)
            , admission_all_covid = admission_child_covid 
                                        + sum(previous_day_admission_adult_covid_suspected 
                                              + previous_day_admission_adult_covid_confirmed
                                              , na.rm = TRUE)
            , non_child_covid_utilization = (inpatient_all_beds_used - total_child_beds_used_covid)/total_all_beds
            , total_utilization = (inpatient_all_beds_used)/total_all_beds
            , .groups = 'keep')

# moving average of the variables above
state_ma_ch <- state_d_ch[order(state_d_ch$hospital_name, state_d_ch$collection_date),] %>%
  group_by(hospital_name) %>%
  mutate(
            inpatient_all_beds = rollmean(inpatient_all_beds, k=7, fill=NA)
            , icu_all_beds = rollmean(icu_all_beds, k=7, fill=NA)
            , total_all_beds = rollmean(total_all_beds, k=7, fill=NA)
            , total_child_beds = rollmean(total_child_beds, k=7, fill=NA)
            , inpatient_child_beds = rollmean(inpatient_child_beds, k=7, fill=NA)
            , icu_child_beds = rollmean(icu_child_beds, k=7, fill=NA)
            , total_all_beds_used_covid = rollmean(total_all_beds_used_covid, k=7, fill=NA)
            , total_child_beds_used_covid_confirmed = rollmean(total_child_beds_used_covid_confirmed, k=7, fill=NA)
            , total_child_beds_used_covid = rollmean(total_child_beds_used_covid, k=7, fill=NA)
            , admission_all_covid = rollmean(admission_all_covid, k=7, fill=NA)
            , admission_child_covid_confirmed = rollmean(admission_child_covid_confirmed, k=7, fill=NA)
            , admission_child_covid = rollmean(admission_child_covid, k=7, fill=NA)
            , non_child_covid_utilization = rollmean(non_child_covid_utilization, k=7, fill=NA)
            , total_utilization = rollmean(total_utilization, k=7, fill=NA)
            ) 

state_ch <- state_ma_ch %>%
  pivot_longer(cols = -c(hospital_name, collection_date)
               , names_pattern = "(...*)_(...*)_beds", names_to = c("bedtype", "scope")) %>%
  mutate(bedtype=ifelse(bedtype=="", "value", bedtype)) 
# %>%  pivot_wider(names_from = 'scope', values_from = 'value')
```




* POWER POINT Graph * 
```{r, fig.height=4}
# New England Children's Hospitals: 
ne_chs <- c("CONNECTICUT CHILDRENS MEDICAL CENTER", "BOSTON CHILDREN'S HOSPITAL", "FRANCISCAN CHILDREN'S HOSPITAL & REHAB CENTER")
# "SHRINERS' HOSPITAL FOR CHILDREN - BOSTON, THE" - For Total all beds they're reporting the beds used, therefor the beds used = total beds
# "SHRINERS' HOSPITAL FOR CHILDREN (THE)" - Only for orthopedics


state_ma_ch %>% 
  filter( hospital_name %in% ne_chs & collection_date > '2020-09-01') %>%
  mutate(r1 = total_utilization) %>%
  mutate(r2 = non_child_covid_utilization) %>%
  ggplot() +  
  facet_wrap(~ hospital_name, ncol = 2) +
  geom_ribbon(aes(ymin = r2, ymax = r1, x=collection_date), fill = "blue") + 
  geom_line(aes(x = collection_date, y = r1, color = "Total Utilization")) +
  geom_line(aes(x = collection_date, y = r2, color = "Non-COVID-19 Utilization")) +
  labs(title = "Children's Hospital Utilization, 7-day moving average", color = "", x = NULL, y = NULL) +
  scale_x_date(breaks = seq.Date(as.Date('2020-11-01'), as.Date('2022-01-01'), "2 month"), date_labels =  "%b %d %Y") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1))
```

```{r}
state_ma_ch %>% 
  filter( hospital_name %in% ne_chs & collection_date %in% c(as.Date('2022-01-05'), as.Date('2022-01-12'))) %>%
  mutate(r1 = total_utilization*100) %>%
  mutate(r2 = non_child_covid_utilization*100) %>%
  mutate(diff = r1-r2) %>%
  filter(!is.na(diff)) %>%
  select(
    hospital_name
    , collection_date
    , r1
    , r2
    , diff
  )
```




Join HHS and Census Data
```{r}
state_ma <- state_ma %>%
  full_join(df_2019child_counts, by = "state")
```

* POWER POINT Graph *
Pediatric Admission Incidence per 100,000 Children
```{r, fig.height=4}
state_ma %>%
  filter( state %in% target_states & collection_date > '2020-09-01') %>%
  mutate(r = admission_child_covid/as.numeric(Data)*100000) %>%
  ggplot(aes(x = collection_date, y = r)) + #, color = state
    facet_wrap(~ state, ncol = 2, scales = 'free_y') + 
    # geom_line(color = "blue", size = 1) +
    geom_point(color = "gray55", size = 1) +
    geom_smooth(color = "blue", size = 1, span = 0.05, se = FALSE) +
    labs(x = NULL, y = 'COVID-19 Pediatric Hospitalizations', title = 'COVID-19 Daily Hospital Admissions per 100K Children,\n 7-day moving average') +
    # scale_x_date(date_breaks = "2 month", date_labels =  "%b-%d-%Y")
    scale_x_date(breaks = seq.Date(as.Date('2020-11-01'), as.Date('2022-01-01'), "2 month"), date_labels =  "%b %d %Y") +
    expand_limits(y = 2) #    scale_y_continuous(expand = c(1, 2))
```
## Slide - R1 and national 7-day averages
```{r, fig.height=4}
col_date <- '2022-01-12'
state_ma %>%
  filter( collection_date > col_date) %>%
  ungroup() %>% group_by(collection_date) %>%
  summarise(US_7DAYAVG = mean(admission_child_covid/as.numeric(Data)*1e5, na.rm = TRUE))
```
```{r, fig.height=4}
state_ma %>%
  filter( state %in% target_states & collection_date > col_date) %>%
  ungroup() %>% group_by(collection_date) %>%
  summarise(R1_7DAYAVERAGE = mean(admission_child_covid/as.numeric(Data)*1e5, na.rm = TRUE))
```

```{r, fig.height=4}
state_ma %>%
  filter( state %in% target_states & collection_date == col_date) %>%
  ungroup() %>% 
  mutate(r = admission_child_covid/as.numeric(Data)*1e5) %>%
  select(collection_date, state, r)
```


Join HHS and Census Data
```{r}
region_ma <- region_ma %>%
  full_join(df_2019child_counts_r, by = c("fema_region" = "region"))
```

* POWER POINT Stat*
Pediatric Case Incidence per 100,000 Children, Regionally
```{r, fig.height=4}
region_ma %>%
  filter(fema_region %in% c("Region 1", "Region 4", "Region 6", "Region 9") & collection_date > '2021-12-21') %>%
  mutate(r = (admission_child_covid/child_population)*100000) #%>%
  #ggplot(aes(x = collection_date, y = r)) + #, color = fema_region
  #  facet_wrap(~ fema_region, ncol = 2) +     #, scales = 'free_y'
  #  geom_line(color = "blue", size = 1) +
  #  labs(x = NULL, y = 'COVID-19 Pediatric Hospitalization', title = 'COVID-19 Hospitalization per 100,000 Children,\n 7-day moving average') +
  #  expand_limits(y = 0)
```


Aggregate by Date
```{r}
us_d <- facility %>%
  group_by(collection_date) %>%
  summarise(
            total_child_beds_used_covid = sum(total_pediatric_patients_hospitalized_confirmed_and_suspected_covid, na.rm = TRUE)
            , total_all_beds_used_covid = total_child_beds_used_covid + sum(total_adult_patients_hospitalized_confirmed_and_suspected_covid, na.rm = TRUE)
            , admission_child_covid_confirmed = sum(previous_day_admission_pediatric_covid_confirmed, na.rm = TRUE)
            , admission_child_covid = admission_child_covid_confirmed +
                                            sum(previous_day_admission_pediatric_covid_suspected, na.rm = TRUE)
            , admission_all_covid = admission_child_covid 
                                        + sum(previous_day_admission_adult_covid_suspected 
                                              + previous_day_admission_adult_covid_confirmed
                                              , na.rm = TRUE)
            , .groups = 'keep')

# moving average of the variables above
us_ma <- us_d[order(us_d$collection_date),] %>%
  ungroup() %>%
  mutate(
            total_all_beds_used_covid = rollmean(total_all_beds_used_covid, k=7, fill=NA)
            , total_child_beds_used_covid = rollmean(total_child_beds_used_covid, k=7, fill=NA)
            , admission_all_covid = rollmean(admission_all_covid, k=7, fill=NA)
            , admission_child_covid_confirmed = rollmean(admission_child_covid_confirmed, k=7, fill=NA)
            , admission_child_covid = rollmean(admission_child_covid, k=7, fill=NA)
            ) 

```


```{r}
us_ma %>%
  filter(collection_date > '2021-12-21') %>%
  mutate(r = (admission_child_covid / 73000000)*100000)
```

See the latest report date
```{r}
max(facility$collection_date)
#colnames(facility)
```



Exploratory Charts
Pediatric COVID hospitalizations - State Level
```{r, fig.width=8}
state_ma %>%
  filter( state %in% target_states & collection_date > '2020-09-01') %>%
  mutate(r = admission_child_covid) %>%
  ggplot(aes(x = collection_date, y = r)) + #, color = state
    facet_wrap(~ state, ncol = 3) + #scales = 'free_y' - For different axes
    geom_line(color = "blue", size = 1) +
    labs(x = NULL, y = 'Hospitalized Cases', title = 'Hospitalized Pediatric COVID-19 Patients \n(Confirmed + Suspected), 7-day moving average') +
    expand_limits(y = 0)
```



Ratio of pediatric COVID hospitalizations to overall COVID hospitalizations
```{r, fig.width=8}
state_ma %>%
  filter( state %in% target_states & collection_date > '2020-09-01') %>%
  mutate(r = (total_child_beds_used_covid / total_all_beds_used_covid)*100) %>%
  ggplot(aes(x = collection_date, y = r)) + #, color = state
    facet_wrap(~ state, ncol = 3) + #scales = 'free_y' - For different axes
    geom_line(color = "blue", size = 1) +
    labs(x = NULL, y = 'Pediatric Hospital Cases/Total Hospital Cases (%)', title = 'Percent Pediatric Hospitalized COVID-19 Patients \n(Confirmed + Suspected), 7-day moving average') +
    expand_limits(y = 0)
```


Ratio of pediatric COVID hospitalizations to overall COVID hospitalizations by Region
```{r, fig.height= 4}
region_ma %>%
  filter(fema_region %in% c("Region 1", "Region 4", "Region 6", "Region 9") & collection_date > '2020-09-01') %>%
  mutate(r = (total_child_beds_used_covid / total_all_beds_used_covid)*100) %>%
  ggplot(aes(x = collection_date, y = r)) + #, color = fema_region
    facet_wrap(~ fema_region, ncol = 2) +     #, scales = 'free_y'
    geom_line(color = "blue", size = 1) +
    labs(x = NULL, y = 'Percent (%)', title = 'Percent Pediatric/Total patients with COVID-19 \n(Confirmed + Suspected), 7-day moving average') +
    expand_limits(y = 0)
```




Aggregate by Hospital for Region 1
```{r}
hospital_d <- facility %>%
  filter(state %in% target_states) %>%
  group_by(hospital_name, collection_date) %>%
  summarise(
            total_all_beds = sum(total_beds, na.rm = TRUE)
            , inpatient_all_beds = sum(inpatient_beds, na.rm = TRUE)
            , icu_all_beds = sum(total_icu_beds, na.rm = TRUE)
            , total_child_beds = total_all_beds - sum(all_adult_hospital_beds, na.rm = TRUE)
            , inpatient_child_beds = inpatient_all_beds - sum(all_adult_hospital_inpatient_beds, na.rm = TRUE)
            , icu_child_beds = icu_all_beds - sum(total_staffed_adult_icu_beds, na.rm = TRUE)
            , inpatient_all_beds_used = sum(inpatient_beds_used, na.rm = TRUE)
            , inpatient_child_beds_used = inpatient_all_beds_used - sum(all_adult_hospital_inpatient_bed_occupied, na.rm = TRUE)
            , icu_all_beds_used = sum(icu_beds_used, na.rm = TRUE)
            , icu_child_beds_used = icu_all_beds_used - sum(staffed_adult_icu_bed_occupancy, na.rm = TRUE)
            , total_child_beds_used_covid_confirmed = sum(total_pediatric_patients_hospitalized_confirmed_covid, na.rm = TRUE)
            , total_child_beds_used_covid = sum(total_pediatric_patients_hospitalized_confirmed_and_suspected_covid, na.rm = TRUE)
            , total_all_beds_used_covid = total_child_beds_used_covid + sum(total_adult_patients_hospitalized_confirmed_and_suspected_covid, na.rm = TRUE)
            , admission_child_covid_confirmed = sum(previous_day_admission_pediatric_covid_confirmed, na.rm = TRUE)
            , admission_child_covid = admission_child_covid_confirmed +
                                            sum(previous_day_admission_pediatric_covid_suspected, na.rm = TRUE)
            , admission_all_covid = admission_child_covid 
                                        + sum(previous_day_admission_adult_covid_suspected 
                                              + previous_day_admission_adult_covid_confirmed
                                              , na.rm = TRUE)
            , non_child_covid_utilization = ((icu_all_beds_used + inpatient_all_beds_used) - total_child_beds_used_covid)/total_all_beds
            , total_utilization = (icu_all_beds_used + inpatient_all_beds_used)/total_all_beds
            , total_deaths = sum(deaths_covid, na.rm = TRUE)
            , critical_staffing_shortage_today = critical_staffing_shortage_today
            , critical_staffing_shortage_anticipated_within_week = critical_staffing_shortage_anticipated_within_week
            , .groups = 'keep')
```

```{r}
# moving average of the variables above
hospital_ma <- hospital_d[order(hospital_d$hospital_name, hospital_d$collection_date),] %>%
  group_by(hospital_name) %>%
  mutate(
            total_all_beds = rollmean(total_all_beds, k=7, fill=NA)
            , inpatient_all_beds = rollmean(inpatient_all_beds, k=7, fill=NA)
            , icu_all_beds = rollmean(icu_all_beds, k=7, fill=NA)
            , total_child_beds = rollmean(total_child_beds, k=7, fill=NA)
            , inpatient_child_beds = rollmean(inpatient_child_beds, k=7, fill=NA)
            , icu_child_beds = rollmean(icu_child_beds, k=7, fill=NA)
            , total_all_beds_used_covid = rollmean(total_all_beds_used_covid, k=7, fill=NA)
            , total_child_beds_used_covid_confirmed = rollmean(total_child_beds_used_covid_confirmed, k=7, fill=NA)
            , total_child_beds_used_covid = rollmean(total_child_beds_used_covid, k=7, fill=NA)
            , admission_all_covid = rollmean(admission_all_covid, k=7, fill=NA)
            , admission_child_covid_confirmed = rollmean(admission_child_covid_confirmed, k=7, fill=NA)
            , admission_child_covid = rollmean(admission_child_covid, k=7, fill=NA)
            , non_child_covid_utilization = rollmean(non_child_covid_utilization, k=7, fill=NA)
            , total_utilization = rollmean(total_utilization, k=7, fill=NA)
            , total_deaths = rollmean(total_deaths, k=7, fill=NA)
            ) 

hospital <- hospital_ma %>%
  pivot_longer(cols = -c(hospital_name, collection_date)
               , names_pattern = "(...*)_(...*)_beds", names_to = c("bedtype", "scope")) %>%
  mutate(bedtype=ifelse(bedtype=="", "value", bedtype)) 
# %>%  pivot_wider(names_from = 'scope', values_from = 'value')
```




Critical Staffing shortage
```{r}
temp <- hospital_d %>%
   group_by(collection_date) %>%
   summarise(
     critical_staffing_shortage_today = sum(as.logical(critical_staffing_shortage_today), na.rm = TRUE)
   )

temp %>%
  filter(collection_date > '2020-09-01') %>%
  mutate(r = critical_staffing_shortage_today/length(unique(hospital_d$hospital_name))*100) %>%
  ggplot(aes(x = collection_date, y = r)) + #, color = fema_region
    #facet_wrap(~ hospital_name, ncol = 2) +     #, scales = 'free_y'
    geom_line() +
    labs(x = NULL, y = 'Percent of Hospitals (%)', title = 'Region 1 Percent of Hospitals Reporting \nCritical Staffing Shortages') +
    expand_limits(y = 0)

print(length(unique(hospital_d$hospital_name)))
```





Hospital Ped Covid Admissions
```{r, fig.height= 10}
hospital_ma %>%
  filter(collection_date > '2020-09-01') %>%
  mutate(r = admission_child_covid) %>%
  ggplot(aes(x = collection_date, y = r, color = hospital_name)) + #, color = fema_region
    #facet_wrap(~ hospital_name, ncol = 2) +     #, scales = 'free_y'
    geom_line() +
    labs(x = NULL, y = 'Admissions', title = 'Pediatric Daily Admissions for COVID-19 \n(Confirmed + Suspected), 7-day moving average') +
    expand_limits(y = 0)
```



Hospital Covid Deaths
```{r, fig.height= 10}
hospital_ma %>%
  filter(collection_date > '2020-09-01') %>%
  mutate(r = total_deaths) %>%
  ggplot(aes(x = collection_date, y = r, color = hospital_name)) + #, color = fema_region
    #facet_wrap(~ hospital_name, ncol = 2) +     #, scales = 'free_y'
    geom_line() +
    labs(x = NULL, y = 'Deaths', title = 'Daily Deaths from COVID-19 \n(Confirmed + Suspected), 7-day moving average') +
    expand_limits(y = 0)
```






