---
title: "R Notebook"
output: html_notebook
---

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

theme_set( theme_bw() + theme( legend.position="bottom" ) +
             theme( text = element_text(size=14) ) )
```

load data from HHS - COVID-19 Reported Patient Impact and Hospital Capacity by Facility
https://healthdata.gov/Hospital/COVID-19-Reported-Patient-Impact-and-Hospital-Capa/anag-cw7u
The following dataset provides facility-level data for hospital utilization aggregated on a weekly basis (Friday to Thursday). These are derived from reports with facility-level granularity across two main sources: (1) HHS TeleTracking, and (2) reporting provided directly to HHS Protect by state/territorial health departments on behalf of their healthcare facilities.

```{r}
facility <- read.csv('data/COVID-19_Reported_Patient_Impact_and_Hospital_Capacity_by_Facility.csv'
                   , na.strings = '-999999') %>%
  mutate(collection_week = as.Date(collection_week))
```

```{r}
sample(unique(facility$hospital_pk), size = 20)
```

Plot timeseries of individual hospital
```{r}
pk <- '110129'
facility %>%
  pivot_longer(c(total_beds_7_day_avg, inpatient_beds_7_day_avg, total_icu_beds_7_day_avg)) %>%
  filter( hospital_pk == pk) %>%
  ggplot(aes(x = collection_week, y = value, color = name)) +
    geom_jitter(height = 0)
```
```{r}
summary(facility %>%
  select(hospital_pk, collection_week, total_beds_7_day_avg, inpatient_beds_7_day_avg, total_icu_beds_7_day_avg)
  )
```

Aggregate by state
```{r}
state <- facility %>%
  group_by(state, collection_week) %>%
  summarise(
            total_all_beds = sum(total_beds_7_day_avg, na.rm = TRUE)
            , inpatient_all_beds = sum(inpatient_beds_7_day_avg, na.rm = TRUE)
            , icu_all_beds = sum(total_icu_beds_7_day_avg, na.rm = TRUE)
            , total_child_beds = total_all_beds - sum(all_adult_hospital_beds_7_day_avg, na.rm = TRUE)
            , inpatient_child_beds = inpatient_all_beds - sum(all_adult_hospital_inpatient_beds_7_day_avg, na.rm = TRUE)
            , icu_child_beds = icu_all_beds - sum(total_staffed_adult_icu_beds_7_day_avg, na.rm = TRUE)
#            , previous_day_admission_pediatric_covid_confirmed_7_day_sum = 
#            , previous_day_admission_pediatric_covid_confirmed_7_day_coverage = 
#            , previous_day_admission_pediatric_covid_suspected_7_day_coverage
            , .groups = 'keep') %>%
  pivot_longer(cols = -c(state, collection_week)
               , names_pattern = "(...*)_(...*)_beds", names_to = c("bedtype", "scope")) %>%
  mutate(bedtype=ifelse(bedtype=="", "value", bedtype)) 
# %>%  pivot_wider(names_from = 'scope', values_from = 'value')
```

Plot timeseries
```{r}
st <- 'CA'
state %>%
  filter( state == st) %>%
  ggplot(aes(x = collection_week, y = value, color = bedtype)) +
    facet_wrap(~ scope, ncol = 1, scales = 'free_y') +
    geom_line() +
    expand_limits(y = 0)
```

